---
title: "PASH Analysis"
author: "Seraphina Shi"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
setwd("/Users/seraphinashi/Desktop/Projects/PAHS/Xuanwei Study coal and DNA metylation study/Seraphina's folder/PAHS-scripts")

plotFolder <- "../images/03_analysis_v3/"
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=plotFolder,
  cache.path=".cache/",
  duplicate.label="allow"
)

```

```{r}
source('shared_commands.R')
```

```{r}
df <- read.csv("../data/data_cleaned.csv") %>% select(-X)
```

```{r get_5MC_data}
library(readxl)
LEX_5MC_all <- read_excel("../data/new_exposure_files/LEX_5MC_all.xlsx")
# View(LEX_5MC_all)
LEX_5MC_all <- LEX_5MC_all %>% rename("SbjctD" = "subject_id", 
                                      "Visit" = "visit",
                                      "cur_5mc" = "sum.X5.methylchrysene_conc_final.imputed", 
                                      "cum_5mc" = "cumlife_5mc",
                                      "bir_5mc" = "birthyr_5mc",
                                      "cur_5mc_measured" = "_5_methylchrysene_conc_measured")


df <- merge(df, LEX_5MC_all %>% select(SbjctD, cur_5mc, cum_5mc, bir_5mc) %>% unique(), by = "SbjctD", all.x=T)
df <- merge(df, LEX_5MC_all %>% select(SbjctD, Visit, cur_5mc_measured) %>% unique(), by = c("SbjctD","Visit"), all.x=T)

var_name_list$PCH_5MC <- c("cur_5mc", "cum_5mc", "bir_5mc", "cur_5mc_measured")
```

```{r}

for (var in unname(unlist(var_name_list$cluster_exp))) {
  sd_var <- sd(df[,var], na.rm = T)
  df[,var] = df[,var]/sd_var
}


df$curFuel[df$curFuel != "Smoky"] = "Other"
df$brthFuel[df$brthFuel != "Smoky"] = "Other"
df$cumFuel[df$cumFuel != "Smoky"] = "Other"
```

```{r}
dup_sbjctD <- df$SbjctD[duplicated(df$SbjctD)]
df_sub1 = df[(df$SbjctD %in% dup_sbjctD & df$Visit == 1) | (! df$SbjctD %in% dup_sbjctD), ]
df_subs <- list("df_sub1" = df_sub1,
                "df_sub2" = df[(df$SbjctD %in% dup_sbjctD & df$Visit == 2) | (! df$SbjctD %in% dup_sbjctD),])
```


```{r define_functions_1}
library(lmtest)
preform_lmtest_mix <- function(Y, X, log_trans_X = F, dat = df) {
  tmp <- dat
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")
  
  form_full <- as.formula(paste0(Y," ~ ", X_form, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")))
  lmfit_full <- lm(form_full, data = tmp)
  form_reduced <- as.formula(paste0(Y," ~ ", paste(var_name_list$confounders[-1], collapse  = " + ")))
  lmfit_reduced <- lm(form_reduced, data = tmp)
  
  lrtest <- lrtest(lmfit_full, lmfit_reduced)

  return(round(lrtest$`Pr(>Chisq)`, 4)[2])
}

preform_lmtest_simple <- function(Y, X, dat = df, log_trans_X = F) {
  tmp <- dat
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")
  
  form_full <- as.formula(paste0(Y," ~ ", X_form))
  lmfit_full <- lm(form_full, data = tmp)
  form_reduced <- as.formula(paste0(Y," ~ 1"))
  lmfit_reduced <- lm(form_reduced, data = tmp)
  
  lrtest <- lrtest(lmfit_full, lmfit_reduced)

  return(round(lrtest$`Pr(>Chisq)`, 4)[2])
}


preform_regression <- function(Y, X, log_trans_X = F, adjust_cfd = F, dat = df) {
  tmp <- dat[!is.na(dat[,X]),]
  if(log_trans_X) {
    tmp[,X] <- log(tmp[,X])
  }
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  lmfit <- lm(form, data = tmp)
  coef <- summary(lmfit)$coefficients[2,1]
  coef_text <- as.character(round(coef,4))
  std <- summary(lmfit)$coefficients[2,2]
  std_text <- as.character(round(std,4))
  p_val <- summary(lmfit)$coefficients[2,4]
  sig_star <- ifelse(p_val<0.001, "***",
                ifelse(p_val<0.01, "** ",
                       ifelse(p_val<0.05, "*  ", "   ")))
  sig_level <- ifelse(p_val <= 0.001, "<= 0.001",
                ifelse(p_val <= 0.01, "<= 0.01",
                       ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))
  rlt <- paste0(coef_text, paste0(rep(" ", 7-nchar(coef_text)), collapse = ""),
                sig_star,
                "(", std_text, paste0(rep(" ", 6-nchar(std_text)), collapse = ""), ")")
  ci_l <- coef - 1.96 * std
  ci_u <- coef + 1.96 * std
  return(c(rlt, coef, ci_l, ci_u, p_val, sig_level))
}


preform_regression_exposure_prototype <- function(X, Y, log_trans_X=F) {
  tmp <- df
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")
  
  tmp <- tmp[tmp$Group == "LEX", ]

  form <- as.formula(paste0(Y," ~ ", X_form, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, pval, sig_level, pval_BHadj)
  
  return(coefs)
}


sig_color <- c("<= 0.001" = "#7CAE00", "<= 0.01" = "#00BFC4", "<= 0.05" = "#C77CFF", "> 0.05" = "black")

make_forest_plot <- function(df_plot, exposure){
  df_plot$sig_level <- as.factor(df_plot$sig_level)
  p <- ggplot(data=df_plot, aes(x=EAAs, y=coefficient, ymin=ci_lower, ymax=ci_upper)) +
    geom_point(shape=24, size=2, aes(col = sig_level, fill = sig_level)) +
    geom_errorbar(shape=24, width=0.3, aes(col = sig_level)) +
    geom_hline(yintercept=0, lty=2,color="red") +  # add a dotted line at x=1 after flip
    # ylim(c(-2,2)) +  # not working
    coord_flip() +  # flip coordinates (puts labels on y axis)
    labs(x=expression(log[2]-PFAS))+
    labs(y="Change in year") +
    ggtitle(exposure)+
    theme(plot.title = element_text(size=8.6, hjust = 0.5),
          axis.title.y = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.y = element_text(size=8),
          axis.title.x = element_text(size=8),
          axis.text = element_text(size=7)) +
    scale_color_manual(values = sig_color) +
    scale_fill_manual(values = sig_color)

  return(p)
}

make_forest_plot_lg <- function(df, exposure){
  df$sig_level <- as.factor(df$sig_level)
  p <- ggplot(data=df, aes(x=EAAs, y=coefficient, ymin=ci_lower, ymax=ci_upper)) +
    geom_point(shape=24, size=2, aes(col = sig_level, fill = sig_level)) +
    geom_errorbar(shape=24, width=0.3, aes(col = sig_level)) +
    geom_hline(yintercept=0, lty=2,color="red") +  # add a dotted line at x=1 after flip
    # ylim(c(-2,2)) +  # not working
    coord_flip() +  # flip coordinates (puts labels on y axis)
    labs(x=expression(log[2]-PFAS))+
    labs(y="Change in year") +
    ggtitle(exposure)+
    theme(plot.title = element_text(size=15, hjust = 0.5),
          axis.title.y = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.y = element_text(size=10),
          axis.title.x = element_text(size=10),
          axis.text = element_text(size=10)) +
    scale_color_manual(values = sig_color) +
    scale_fill_manual(values = sig_color)

  return(p)
}


options(scipen = 999)
```

```{r define_functions_2}
library(geepack)

preform_gee <- function(Y, X, log_trans_X = F, adjust_cfd = F, dat = df) {
  tmp <- dat[!is.na(dat[,X]),]
  if(log_trans_X) {
    tmp[,X] <- log(tmp[,X])
  }
  
  if(adjust_cfd){
    tmp <- tmp %>% select(var_name_list$ID,
                          var_name_list$EEAs,
                          var_name_list$confounders,
                          unname(X)) %>% na.omit()
  }
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  X_form <- ifelse(adjust_cfd, 
                 paste0(X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(X))

  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") # This says all pairs of responses within a subject are equally correlated
  coef <- summary(geefit)$coefficients[2,1]
  coef_text <- as.character(round(coef,4))
  std <- summary(geefit)$coefficients[2,2]
  std_text <- as.character(round(std,4))
  p_val <- summary(geefit)$coefficients[2,4]
  sig_star <- ifelse(p_val<0.001, "***",
                ifelse(p_val<0.01, "** ",
                       ifelse(p_val<0.05, "*  ", "   ")))
  sig_level <- ifelse(p_val <= 0.001, "<= 0.001",
                ifelse(p_val <= 0.01, "<= 0.01",
                       ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))
  # rlt <- paste0(coef_text, paste0(rep(" ", 7-nchar(coef_text)), collapse = ""),
  #               sig_star,
  #               "(", std_text, paste0(rep(" ", 6-nchar(std_text)), collapse = ""), ")")
  ci_l <- coef - 1.96 * std
  ci_u <- coef + 1.96 * std
  return(c(coef, std, ci_l, ci_u, p_val, sig_level))
}


perform_group_gee <- function(X, adjust_cfd = F, var_name = NA){
  plot_df_list <- list()
  lm_coef_ambient_plot_list <- list()
  for (j in 1:length(X)) {
    exp <- X[j]
    plot_df <- data.frame(matrix(NA, ncol = 6))
    for(i in 1:length(var_name_list$EEAs)){
      rlts = preform_gee(Y = var_name_list$EEAs[i], 
                                X = exp, 
                                adjust_cfd = adjust_cfd, 
                                dat=df)
      plot_df[i,] <- rlts
    }
    colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
    rownames(plot_df) <- var_name_list$EEAs
    plot_df$EAAs <- EAAs_names
    plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
  
    plot_df$coefficient = as.numeric(plot_df$coefficient)
    plot_df$ci_lower = as.numeric(plot_df$ci_lower)
    plot_df$ci_upper = as.numeric(plot_df$ci_upper)
  
    if(sum(is.na(var_name))>0){
      lm_coef_ambient_plot <- make_forest_plot(plot_df, exp)
    } else {
      lm_coef_ambient_plot <- make_forest_plot(plot_df, var_name[j])
    }
    
    plot_df_list[[exp]] <- plot_df
    lm_coef_ambient_plot_list[[exp]] = lm_coef_ambient_plot
  }
  return(list(plot_df_list, lm_coef_ambient_plot_list))
}



preform_gee_exposure_prototype <- function(X, Y, adjust_cfd = T, log_trans_X=F) {
  tmp <- df
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")

  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X_form, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~", X_form))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") # This says all pairs of responses within a subject are equally correlated


  coefs <- as.data.frame(summary(geefit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")

  coefs <- coefs %>%
    select(coefficient, std, pval) %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, pval, sig_level)


  colnames(coefs) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  
  return(coefs)
}

```






In this file, we presented the results from analyses of exposures against computed epigenetic age accelerations (EAA) calculated using DNA methylation data by different methods. For each exposure, we conducted both primary analyses (likelihood ratio tests and generalized estimating equations (GEE), adjusted for confounders) and sensitivity analyses (likelihood ratio tests and/or generalized estimating equations (GEE) limited to certain fuel users, not adjusted for confounders). In addition to what's included in the analysis plan, we also analyzed ambient and urinary exposures. 


# 1.1. Description of study population

There are `r nrow(df)` visits with corresponding epigenetic ages available among `r length(unique(df$SbjctD))` female subjects. For these `r length(unique(df$SbjctD))` subjects, `r length(unique(df$SbjctD)) - length(dup_sbjctD)` have been visited once and `r length(dup_sbjctD)` have been visited twice. 

The following tables summarize all the information of the first visit of these `r length(unique(df$SbjctD))` subjects.

## Baseline characteristics (confounders)
```{r}
library(table1)

# df_sub1 %>% select(var_name_list$confounders) %>% mutate(ses = as.factor(ses)) %>%
#   tbl_summary(statistic=list(all_continuous() ~ "{mean} ({sd})", 
#                              all_categorical() ~ "{n} / {N} ({p}%)"),
#               missing_text = "(Missing)")

table1(~ Age + BMI + factor(county) + factor(ses)  + factor(edu), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```


## Epigenetic ages
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$clocks, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Epigenetic ages accelarations
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$EEAs, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Fuel/stove type exposures
```{r}
table1(as.formula(paste(" ~ ", paste(paste0("factor(", c(var_name_list$fuel_exp, var_name_list$stove_exp),")"), 
                                     collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## 5MC exposures
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```


```{r}
print("Pearson pair-wise correlation:")
cor(df_sub1 %>%  select(var_name_list$PCH_5MC) %>% na.omit(), 
    method = "pearson")
```
```{r}
print("Spearman pair-wise correlation:")
cor(df_sub1 %>%  select(var_name_list$PCH_5MC) %>% na.omit(), 
    method = "spearman")
```

## Cluster-based exposures
### clusCUR6
Clusters based on model-based exposure estimates at or shortly before the visit
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUR6, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

### clusCHLD5
Clusters based on model-based exposure estimates accrued before age 18 
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCHLD5, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

### clusCUM6
Clusters based on model-based lifetime exposure estimates 
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUM6, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

### clusMEAS6
Clusters based on pollutant measurements
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusMEAS6, collapse = " + "))), 
       data = df_sub1)
```

### clusURI5
Clusters based on urinary biomarkers
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusURI5, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```



## Ambient exposures
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$ambient_exp, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Urinary biomarkers
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$urinary_exp, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```


# 1.2. EAA ~ confounders
```{r gee_EAA_confounders}
for(i in 1:length(var_name_list$EEAs)){
    cat("For", var_name_list$EEAs[i], ":")
    rlts = preform_gee_exposure_prototype(X = var_name_list$confounders,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = F)
    rlts$coefficient = round(as.numeric(rlts$coefficient), 4)
    rlts$std = round(as.numeric(rlts$std), 4)
    rlts$ci_lower = round(as.numeric(rlts$ci_lower), 4)
    rlts$ci_upper = round(as.numeric(rlts$ci_upper), 4)
    rlts$p_val = round(as.numeric(rlts$p_val), 4)
    print(rlts)
}
```


# 1.3. 5MC ~ confounders
```{r gee_5MC_confounders}
for(i in 1:length(var_name_list$PCH_5MC)){
    cat("For", var_name_list$PCH_5MC[i], ":")
    rlts = preform_gee_exposure_prototype(X = var_name_list$confounders,
                                          Y = var_name_list$PCH_5MC[i],
                                          adjust_cfd = F)
    rlts$coefficient = round(as.numeric(rlts$coefficient), 4)
    rlts$std = round(as.numeric(rlts$std), 4)
    rlts$ci_lower = round(as.numeric(rlts$ci_lower), 4)
    rlts$ci_upper = round(as.numeric(rlts$ci_upper), 4)
    rlts$p_val = round(as.numeric(rlts$p_val), 4)

    print(rlts)
}
```
# 1.4. clusCUR6 ~ confounders
```{r gee_clusCUR6_confounders}
for(i in 1:length(var_name_list$cluster_exp$clusCUR6)){
    cat("For", var_name_list$cluster_exp$clusCUR6[i], ":")
    rlts = preform_gee_exposure_prototype(X = var_name_list$confounders,
                                          Y = var_name_list$cluster_exp$clusCUR6[i],
                                          adjust_cfd = F)
    rlts$coefficient = round(as.numeric(rlts$coefficient), 4)
    rlts$std = round(as.numeric(rlts$std), 4)
    rlts$ci_lower = round(as.numeric(rlts$ci_lower), 4)
    rlts$ci_upper = round(as.numeric(rlts$ci_upper), 4)
    rlts$p_val = round(as.numeric(rlts$p_val), 4)

    print(rlts)
}
```

# 2.1. Current (self-reported) fuel type 
The numbers of observations with each current fuel type:
```{r}
table(df$curFuel)
```

## Primary analysis
Investigate the association with current (self-reported) fuel type in the LEX study participants, adjusting for known confounders. The reference group for this analysis would be the smoky coal users. This would be a categorical analysis, and the results would be a p-value from the likelihood ratio (LR) test of a confounder-only model to a model including the exposure variables, as well as p-values for the contrast of each category of coal use (smokeless coal or plant/wood) to that of smoky coal. FDR correction should be used separately for each of these sets. The main interest would be in the coal-specific findings and perhaps less so in the results from the LR test. 

### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$


Results:
```{r}
preform_gee_fuel <- function(Y, X, adjust_cfd = F) {
  tmp <- df[!is.na(df[,X]),]
  
  tmp[,X] = factor(tmp[,X], levels = c("Other","Smoky"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Other (reference/intercept)"
  rownames(coefs)[grepl("Smoky", rownames(coefs))] = "Smoky"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_curFuel_with_confounders, fig.width=4, fig.height=3}
coefs_1 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_fuel(X = "curFuel", Y = var_name_list$EEAs[i], adjust_cfd = T)
    coefs_1[i,] <- rlts[2,]
}

colnames(coefs_1) = colnames(rlts)
rownames(coefs_1) = var_name_list$EEAs
coefs_1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

cat("Average EAA difference of current Smoky fuel type compared to Other fuel types")
p1 <- make_forest_plot_lg(coefs_1, 
                          "")
p1
```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to other fuel yupes:")
print(coefs_1 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 4)) %>%
        select(-EAAs))
```


## Sensitivity analyses
### GEE (no confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \epsilon
\end{aligned}
$$


Results:

```{r gee_curFuel_no_confounders, fig.width=4, fig.height=3}
coefs_1_simple <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_fuel(X = "curFuel", Y = var_name_list$EEAs[i], adjust_cfd = F)
    coefs_1_simple[i,] <- rlts[2,]
}

colnames(coefs_1_simple) = colnames(rlts)
rownames(coefs_1_simple) = var_name_list$EEAs
coefs_1_simple$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

cat("Average EAA difference of current Smoky fuel type compared to Other fuel types")
p1_simple <- make_forest_plot_lg(coefs_1_simple, 
                          "")
p1_simple
```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to other fuel yupes:")
print(coefs_1_simple %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 4)) %>%
        select(-EAAs))
```

# 2.2. Childhood  (self-reported) fuel type 
The numbers of observations with each current fuel type:
```{r}
table(df$brthFuel)
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$


Results:
```{r gee_brthFuel_with_confounders, fig.width=4, fig.height=3}
coefs_2 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_fuel(X = "brthFuel", Y = var_name_list$EEAs[i], adjust_cfd = T)
    coefs_2[i,] <- rlts[2,]
}

colnames(coefs_2) = colnames(rlts)
rownames(coefs_2) = var_name_list$EEAs
coefs_2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

cat("Average EAA difference of birth Smoky fuel type compared to Other fuel types")
p3 <- make_forest_plot_lg(coefs_2, 
                          "")
p3
```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to other fuel yupes:")
print(coefs_2 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 4)) %>%
        select(-EAAs))
```


## Sensitivity analyses
### GEE (no confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \epsilon
\end{aligned}
$$


Results:
```{r gee_brthFuel_no_confounders, fig.width=4, fig.height=3}
coefs_2_simple <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_fuel(X = "brthFuel", Y = var_name_list$EEAs[i], adjust_cfd = T)
    coefs_2_simple[i,] <- rlts[2,]
}

colnames(coefs_2_simple) = colnames(rlts)
rownames(coefs_2_simple) = var_name_list$EEAs
coefs_2_simple$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

cat("Average EAA difference of birth Smoky fuel type compared to Other fuel types")
p3_simple <- make_forest_plot_lg(coefs_2_simple, 
                          "")
p3_simple
```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to other fuel yupes:")
print(coefs_2_simple %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 4)) %>%
        select(-EAAs))
```

# 2.2. Cumulative lifetime (self-reported) fuel type 
The numbers of observations with each cumulative lifetime fuel type:
```{r}
table(df$cumFuel)
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$


Results:
```{r gee_cumFuel_with_confounders, fig.width=4, fig.height=3}
coefs_3 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_fuel(X = "cumFuel", Y = var_name_list$EEAs[i], adjust_cfd = T)
    coefs_3[i,] <- rlts[2,]
}

colnames(coefs_3) = colnames(rlts)
rownames(coefs_3) = var_name_list$EEAs
coefs_3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

cat("Average EAA difference of cumulative Smoky fuel type compared to Other fuel types")
p2 <- make_forest_plot_lg(coefs_3, 
                          "")
p2
```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to other fuel yupes:")
print(coefs_3 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 4)) %>%
        select(-EAAs))
```


## Sensitivity analyses
### GEE (no confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \epsilon
\end{aligned}
$$


Results:
```{r gee_cumFuel_no_confounders, fig.width=4, fig.height=3}
coefs_3_simple <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_fuel(X = "cumFuel", Y = var_name_list$EEAs[i], adjust_cfd = F)
    coefs_3_simple[i,] <- rlts[2,]
}

colnames(coefs_3_simple) = colnames(rlts)
rownames(coefs_3_simple) = var_name_list$EEAs
coefs_3_simple$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

cat("Average EAA difference of cumulative Smoky fuel type compared to Other fuel types")
p2_simple <- make_forest_plot_lg(coefs_3_simple, 
                          "")
p2_simple
```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to other fuel yupes:")
print(coefs_3_simple %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 4)) %>%
        select(-EAAs))
```




# 2.0. All plots together
```{r gee_FuelTypes_3periods, fig.width=8, fig.height=3}
coefs_1$exposure_time = "Current Exposure"
coefs_2$exposure_time = "Childhood Exposure"
coefs_3$exposure_time = "Cumulative Exposure"
coefs <- rbind(coefs_1, coefs_2, coefs_3)
coefs$exposure_time = factor(coefs$exposure_time, levels = c("Current Exposure","Childhood Exposure", "Cumulative Exposure"))
make_forest_plot_lg(coefs, "Average EAA Difference of Smoky Fuel Type Compared to Other Fuel Type") + 
  facet_wrap(~ exposure_time)
```






# 3.1. Clusters based on model-based exposure estimates at or shortly before the visit (clusCUR6)
The file “LEX_clusCUR6.csv” has information on current pollutant exposures, obtained for the 2 years preceding the visit. To reduce multi-collinearity between exposures, exposure prototypes were derived based on hierarchical cluster analysis in combination followed by principal components analysis. These estimates are available for 6 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:

CUR6_BC_PAH6 – Black carbon (BC) and 6 PAHs   
CUR6_PAH31 – a large cluster of 31 PAHs  
CUR6_NkF – NkF only  
CUR6_PM_RET – Particulate matter (PM) and retene  
CUR6_NO2 – NO2 only  
CUR6_SO2 – SO2 only    



## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUR6, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis

### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUR6_mix, fig.width=8, fig.height=7}
rlts_gee_1 <- perform_group_gee(X = var_name_list$cluster_exp$clusCUR6, adjust_cfd = T)

for(j in 1:length(rlts_gee_1[[2]])){
  rlts_gee_1[[2]][[j]] <- rlts_gee_1[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee_1[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee_1[[1]])){
  rlts_gee_1[[1]][[j]] <- rlts_gee_1[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee_1[[1]])
```


### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUR6_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUR6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CUR6_BC_PAH6")
p2 <- make_forest_plot_lg(coef2, "CUR6_PAH31")
p3 <- make_forest_plot_lg(coef3, "CUR6_NkF")
p4 <- make_forest_plot_lg(coef4, "CUR6_PM_RET")
p5 <- make_forest_plot_lg(coef5, "CUR6_NO2")
p6 <- make_forest_plot_lg(coef6, "CUR6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUR6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```

## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUR6_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusCUR6, adjust_cfd = F)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee[[1]])
```




### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  &  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUR6_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUR6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = F)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CUR6_BC_PAH6")
p2 <- make_forest_plot_lg(coef2, "CUR6_PAH31")
p3 <- make_forest_plot_lg(coef3, "CUR6_NkF")
p4 <- make_forest_plot_lg(coef4, "CUR6_PM_RET")
p5 <- make_forest_plot_lg(coef5, "CUR6_NO2")
p6 <- make_forest_plot_lg(coef6, "CUR6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUR6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```


# 3.2. Clusters based on model-based exposure estimates accrued before age 18 (clusCHLD5)
The file “LEX_clusCHLD5.csv” has information on estimated pollutant exposures during early childhood. Estimates are available for 5 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:

CHLD5_X7 – a cluster of 7 air pollutants  
CHLD5_X33 – a large cluster of 33 air pollutants  
CHLD5_NkF – NkF only  
CHLD5_NO2 – NO2 only  
CHLD5_SO2 – SO2 only    


## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCHLD5, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCHLD5_mix, fig.width=8, fig.height=7}
rlts_gee_2 <- perform_group_gee(X = var_name_list$cluster_exp$clusCHLD5, adjust_cfd = T)

for(j in 1:length(rlts_gee_2[[2]])){
  rlts_gee_2[[2]][[j]] <- rlts_gee_2[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee_2[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee_2[[1]])){
  rlts_gee_2[[1]][[j]] <- rlts_gee_2[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee_2[[1]])
```




### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X7 +  \beta_2 X33 + \beta_3 NkF + \beta_4 NO2 + \beta_5 SO2  \\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_9 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCHLD5_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCHLD5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CHLD5_X7")
p2 <- make_forest_plot_lg(coef2, "CHLD5_X33")
p3 <- make_forest_plot_lg(coef3, "CHLD5_NkF")
p4 <- make_forest_plot_lg(coef4, "CHLD5_NO2")
p5 <- make_forest_plot_lg(coef5, "CHLD5_SO2")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusCHLD5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```



## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCHLD5_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusCHLD5, adjust_cfd = F)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee[[1]])
```





### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X7 +  \beta_2 X33 + \beta_3 NkF + \beta_4 NO2 + \beta_5 SO2  \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCHLD5_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCHLD5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = F)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CHLD5_X7")
p2 <- make_forest_plot_lg(coef2, "CHLD5_X33")
p3 <- make_forest_plot_lg(coef3, "CHLD5_NkF")
p4 <- make_forest_plot_lg(coef4, "CHLD5_NO2")
p5 <- make_forest_plot_lg(coef5, "CHLD5_SO2")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-4,4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusCHLD5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```




# 3.3. Clusters based on model-based lifetime exposure estimates (clusCUM6)
The file “LEX_clus CUM6.csv” has information on estimated cumulative pollutant exposures during the lifecourse. Estimates are available for 6 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:  

CUM6_BC_NO2_PM – a cluster of BC, NO2, and PM  
CUM6_PAH36 – a large cluster of 36 PAHs  
CUM6_DlP – DlP only  
CUM6_NkF – NkF only  
CUM6_RET – retene only  
CUM6_SO2 – SO2 only   
 

## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUM6, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUM6_mix, fig.width=8, fig.height=7}
rlts_gee_3 <- perform_group_gee(X = var_name_list$cluster_exp$clusCUM6, adjust_cfd = T)

for(j in 1:length(rlts_gee_3[[2]])){
  rlts_gee_3[[2]][[j]] <- rlts_gee_3[[2]][[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = rlts_gee_3[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee_3[[1]])){
  rlts_gee_3[[1]][[j]] <- rlts_gee_3[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee_3[[1]])
```



### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_NO2\_PM +  \beta_2 PAH36 + \beta_3 DlP  + \beta_4 NkF + \beta_5 RET + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUM6_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUM6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CUM6_BC_NO2_PM")
p2 <- make_forest_plot_lg(coef2, "CUM6_PAH36")
p3 <- make_forest_plot_lg(coef3, "CUM6_DlP")
p4 <- make_forest_plot_lg(coef4, "CUM6_NkF")
p5 <- make_forest_plot_lg(coef5, "CUM6_RET")
p6 <- make_forest_plot_lg(coef6, "CUM6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUM6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```


## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUM6_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusCUM6, adjust_cfd = F)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee[[1]])
```



### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_NO2\_PM +  \beta_2 PAH36 + \beta_3 DlP  + \beta_4 NkF + \beta_5 RET + \beta_6 SO2  \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUM6_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUM6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = F)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CUM6_BC_NO2_PM")
p2 <- make_forest_plot_lg(coef2, "CUM6_PAH36")
p3 <- make_forest_plot_lg(coef3, "CUM6_DlP")
p4 <- make_forest_plot_lg(coef4, "CUM6_NkF")
p5 <- make_forest_plot_lg(coef5, "CUM6_RET")
p6 <- make_forest_plot_lg(coef6, "CUM6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUM6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```




# 3.0. All plots together
```{r fig.width=15, fig.height=8}
rlts_gee_1[[1]]$CUR6_BC_PAH6$cluster = "BC_PAH6"
rlts_gee_1[[1]]$CUR6_PAH31$cluster = "PAH31"
rlts_gee_1[[1]]$CUR6_NkF$cluster = "NkF"
rlts_gee_1[[1]]$CUR6_PM_RET$cluster = "PM_RET"
rlts_gee_1[[1]]$CUR6_NO2$cluster = "NO2"
rlts_gee_1[[1]]$CUR6_SO2$cluster = "SO2"

rlts_gee_1_Df <- rbind(rlts_gee_1[[1]]$CUR6_BC_PAH6, rlts_gee_1[[1]]$CUR6_PAH31, 
                       rlts_gee_1[[1]]$CUR6_NkF, rlts_gee_1[[1]]$CUR6_PM_RET,
                       rlts_gee_1[[1]]$CUR6_NO2, rlts_gee_1[[1]]$CUR6_SO2)
rlts_gee_1_Df$exposure_period = "Current Exposure"

rlts_gee_2[[1]]$CHLD5_X7$cluster = "X7"
rlts_gee_2[[1]]$CHLD5_X33$cluster = "X33"
rlts_gee_2[[1]]$CHLD5_NkF$cluster = "NkF"
rlts_gee_2[[1]]$CHLD5_NO2$cluster = "NO2"
rlts_gee_2[[1]]$CHLD5_SO2$cluster = "SO2"
empty_df <- as.data.frame(matrix(NA, nrow = 8, ncol = 6)) 
empty_df$cluster = " "
row.names(empty_df) <- row.names(rlts_gee_2[[1]]$CHLD5_X7)
colnames(empty_df) <- colnames(rlts_gee_2[[1]]$CHLD5_X7)

rlts_gee_2_Df <- rbind(rlts_gee_2[[1]]$CHLD5_X7, rlts_gee_2[[1]]$CHLD5_X33,
                       rlts_gee_2[[1]]$CHLD5_NkF, rlts_gee_2[[1]]$CHLD5_NO2,
                       rlts_gee_2[[1]]$CHLD5_SO2, empty_df)
rlts_gee_2_Df$exposure_period = "Childhood Exposure"


rlts_gee_3[[1]]$CUM6_BC_NO2_PM$cluster = "BC_NO2_PM"
rlts_gee_3[[1]]$CUM6_PAH36$cluster = "PAH36"
rlts_gee_3[[1]]$CUM6_DlP$cluster = "DlP"
rlts_gee_3[[1]]$CUM6_NkF$cluster = "NkF"
rlts_gee_3[[1]]$CUM6_RET$cluster = "RET"
rlts_gee_3[[1]]$CUM6_SO2$cluster = "SO2"

rlts_gee_3_Df <- rbind(rlts_gee_3[[1]]$CUM6_BC_NO2_PM, rlts_gee_3[[1]]$CUM6_PAH36, 
                       rlts_gee_3[[1]]$CUM6_DlP, rlts_gee_3[[1]]$CUM6_NkF,
                       rlts_gee_3[[1]]$CUM6_RET, rlts_gee_3[[1]]$CUM6_SO2)
rlts_gee_3_Df$exposure_period = "Cumulative Exposure"

rlts_gee_Df <- rbind(rlts_gee_1_Df, rlts_gee_2_Df, rlts_gee_3_Df)


rlts_gee_Df$exposure_period = factor(rlts_gee_Df$exposure_period, 
                                     levels = c("Current Exposure","Childhood Exposure", "Cumulative Exposure"))
rlts_gee_Df$cluster = factor(rlts_gee_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36", "DlP", "BC_PAH6", "PAH31", "X7", "X33", 
                                                "PM_RET",  " ", "NkF", "RET","NO2", "SO2"))
# rlts_gee_Df$cluster = factor(rlts_gee_Df$cluster, 
#                                      levels = c("BC_NO2_PM", "PAH36", "DlP", "BC_PAH6", "PAH31", "X7", "X33", 
#                                                 "PM_RET","RET",  "NkF", "NO2", "SO2", " "))
rlts_gee_Df$EAAs <- rep(var_name_list$EEAs, 6+6+6)
```


```{r gee_clusters_3periods, fig.width=15, fig.height=8}
make_forest_plot_lg(rlts_gee_Df, "Expected EAA Change with 1 Unit Increased in each Cluster") + 
  facet_wrap( ~ exposure_period + cluster, nrow=3)
```





# 3.4. Clusters based on pollutant measurements (clusMEAS6)
The file “LEX_clusMEAS6.csv” has information on measured pollutant exposures during each visit. Estimates are available for 6 different prototypes (cluster variables) for a total of 54 subjects and 54 visits. The prototypes are labelled as:   

MEAS6_BC_ PM_RET – a cluster of BC, PM, and retene  
MEAS6_X31 – a large cluster of 31 air pollutants  
MEAS6_X5 – a smaller cluster of 5 air pollutants  
MEAS6_DlP – DlP only  
MEAS6_NkF – NkF only  
MEAS6_ NO2_SO2 – NO2, and SO2    


## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusMEAS6, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusMEAS6_mix, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusMEAS6, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee[[1]])
```



### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusMEAS6_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusMEAS6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "MEAS6_BC_PM_RET")
p2 <- make_forest_plot_lg(coef2, "MEAS6_X31")
p3 <- make_forest_plot_lg(coef3, "MEAS6_X5")
p4 <- make_forest_plot_lg(coef4, "MEAS6_DlP")
p5 <- make_forest_plot_lg(coef5, "MEAS6_NkF")
p6 <- make_forest_plot_lg(coef6, "MEAS6_NO2_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusMEAS6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```


## Sensitivity analyses
### GEE (no confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusMEAS6_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusMEAS6, adjust_cfd = F)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee[[1]])
```




### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusMEAS6_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusMEAS6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = F)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "MEAS6_BC_PM_RET")
p2 <- make_forest_plot_lg(coef2, "MEAS6_X31")
p3 <- make_forest_plot_lg(coef3, "MEAS6_X5")
p4 <- make_forest_plot_lg(coef4, "MEAS6_DlP")
p5 <- make_forest_plot_lg(coef5, "MEAS6_NkF")
p6 <- make_forest_plot_lg(coef6, "MEAS6_NO2_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusMEAS6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```



# 3.5. Clusters based on urinary biomarkers (clusURI5)
The file “LEX_clusURI5.csv” has information on measured urinary biomarkers obtained during each visit. Estimates are available for 5 different prototypes (cluster variables) for a total of 163 subjects and 186 visits. The prototypes are labelled as:  


URI5_NAP_1M_2M – a cluster of Naphthalene, 1Methylnaphthalene, and 2Methylnaphthalene  
URI5_ACE – Acenaphthene only  
URI5_FLU_PHE – Fluoranthene and Phenanthrene_anth  
URI5_PYR – Pyrene only  
URI5_CHR – Baa_Chrysene only  
 



## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusURI5, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusURI5_mix, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusURI5, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured urinary biomarkers"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee[[1]])
```



### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * NAP\_1M\_2M +  \beta_2 ACE + \beta_3 FLU\_PHE + \beta_4 PYR + \beta_5 CHR  \\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_9 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusURI5_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusURI5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "URI5_NAP_1M_2M")
p2 <- make_forest_plot_lg(coef2, "URI5_ACE")
p3 <- make_forest_plot_lg(coef3, "URI5_FLU_PHE")
p4 <- make_forest_plot_lg(coef4, "URI5_PYR")
p5 <- make_forest_plot_lg(coef5, "URI5_CHR")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured urinary biomarkers"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusURI5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```


## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusURI5_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusURI5, adjust_cfd = F)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(rlts_gee[[1]])
```



### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * NAP\_1M\_2M +  \beta_2 ACE + \beta_3 FLU\_PHE + \beta_4 PYR + \beta_5 CHR  \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusURI5_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusURI5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = F)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "URI5_NAP_1M_2M")
p2 <- make_forest_plot_lg(coef2, "URI5_ACE")
p3 <- make_forest_plot_lg(coef3, "URI5_FLU_PHE")
p4 <- make_forest_plot_lg(coef4, "URI5_PYR")
p5 <- make_forest_plot_lg(coef5, "URI5_CHR")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured urinary biomarkers"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusURI5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 4))
}
print(coefs_list)
```



# 4.1. Current exposure to 5MC
## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC[1], collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```
## Primary analysis
### GEE (with confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cur\_5mc\\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cur_5MC, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cur_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Current exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)
plot_df_1 <- plot_df
```





## Sensitivity analysis
### GEE (no confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cur\_5mc + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cur_5MC_simple, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cur_5mc", 
                              adjust_cfd = F, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Current exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)
```





# 4.2. Childhood exposure to 5MC
## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC[3], collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```
## Primary analysis

### GEE (with confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *childhood\_5mc\\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_bir_5MC, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "bir_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Childhood exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)
plot_df_2 <- plot_df
```





## Sensitivity analysis
### GEE (no confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *childhood\_5mc + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cir_5MC_simple, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cur_5mc", 
                              adjust_cfd = F, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Childhood exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)

```



# 4.3. Cumulative exposure to 5MC
## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC[2], collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```
## Primary analysis

### GEE (with confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cum\_5mc\\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cum_5MC, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cum_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Cumulatice exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)

plot_df_3 <- plot_df
```





## Sensitivity analysis
### GEE (no confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cum\_5mc + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cum_5MC_simple, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cum_5mc", 
                              adjust_cfd = F, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Current exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)

```



# 4.0. All plots together
```{r gee_5MC_3Periods, fig.width=8, fig.height=3}
plot_df_1$exposure_time = "Current Exposure"
plot_df_2$exposure_time = "Childhood Exposure"
plot_df_3$exposure_time = "Cumulative Exposure"
plot_df_5mc <- rbind(plot_df_1, plot_df_2, plot_df_3)
plot_df_5mc$exposure_time = factor(plot_df_5mc$exposure_time, levels = c("Current Exposure","Childhood Exposure", "Cumulative Exposure"))
make_forest_plot_lg(plot_df_5mc, "Expected EAA Change with 1 Unit Increased in 5MC Exposure") + 
  facet_wrap(~ exposure_time, scales = 'free_x') 
```
