---
title: "Analysis"
author: "Seraphina Shi"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
setwd("/Users/seraphinashi/Desktop/Projects/PAHS/Xuanwei Study coal and DNA metylation study/Seraphina's folder/PAHS-scripts")

plotFolder <- "../images/03_analysis/"
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=plotFolder,
  cache.path=".cache/",
  duplicate.label="allow"
)

```

```{r}
source('shared_commands.R')
```

```{r}
df <- read.csv("../data/data_cleaned.csv")
```

# Current (self-reported) fuel type 
The numbers of observations with each current fuel type:
```{r}
table(df$curFuel)
```

## Primary analysis
Investigate the association with current (self-reported) fuel type in the LEX study participants, adjusting for known confounders and stove ventilation. The reference group for this analysis would be the smoky coal users. This would be a categorical analysis, and the results would be a p-value from the likelihood ratio (LR) test of a confounder-only model to a model including the exposure variables, as well as p-values for the contrast of each category of coal use (smokeless coal or plant/wood) to that of smoky coal. FDR correction should be used separately for each of these sets. The main interest would be in the coal-specific findings and perhaps less so in the results from the LR test. 

### Likelihood ratio (LR) test (mix model)
Full model: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smokeles}) + \beta_2 * I(\text{Wood_and_or_Plant})   \\
  & + \beta_3 * county + \beta_4 * BMI + \beta_5 * ses + \beta_6 * edu + \beta_7 * curStove + \epsilon
\end{aligned}
$$
Nested model: 
$$
\begin{aligned}
  Y = & \beta_0   \\
  & + \beta_1 * county + \beta_2 * BMI + \beta_3 * ses + \beta_4 * edu + \beta_5 * curStove + \epsilon
\end{aligned}
$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
library(lmtest)
preform_lmtest_mix <- function(Y, X, log_trans_X = F) {
  tmp <- df
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")
  
  form_full <- as.formula(paste0(Y," ~ ", X_form, " + ", paste(var_name_list$confounders[-1], collapse  = " + "), " + curStove"))
  lmfit_full <- lm(form_full, data = tmp)
  form_reduced <- as.formula(paste0(Y," ~ ", paste(var_name_list$confounders[-1], collapse  = " + "), " + curStove"))
  lmfit_reduced <- lm(form_reduced, data = tmp)
  
  lrtest <- lrtest(lmfit_full, lmfit_reduced)

  return(round(lrtest$`Pr(>Chisq)`, 4)[2])
}
```

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_mix(Y = EAAs_vars[i], X = "curFuel")
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```
```{r}
sig_color <- c("<= 0.001" = "#7CAE00", "<= 0.01" = "#00BFC4", "<= 0.05" = "#C77CFF", "> 0.05" = "black")

library('plot.matrix')
library('psych')
par(mar=c(7.1, 4.1, 4.1, 4.1)) # adapt margins
# plot(p_vals_mat, cex=0.5, assignColors())

# x <- runif(100)
plot(p_vals_mat, 
     border=NA,
     axis.row=list(side=2, las=1),
     col = assignColors(c(0:1000/1000), 
                        breaks=c(0, 0.001, 0.01,0.05,1),
                  col = c("#7CAE00", "#00BFC4", "#C77CFF", "gray87")))
# set breaks
# assignColors(x, breaks=15)
# assignColors(x, breaks=c(0,1))

plot(p_vals_mat)

```


### Linear regression 
In the following section, we performed linear regression with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smokeles}) + \beta_2 * I(\text{Wood_and_or_Plant})   \\
  & + \beta_3 * county + \beta_4 * BMI + \beta_5 * ses + \beta_6 * edu + \beta_7 * curStove + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and the reference as the smoky fuel type.

The estimations of $\beta_0$, $\beta_1$ and $\beta_2$ with given $Y$ are shown below. The $\beta_1$ and $\beta_2$ can be interpreted as "the expected change of Y if switching form the smoky fuel type to the given fuel type, while holding other variables constant".


```{r}
preform_regression_curFuel_mix <- function(Y) {
  X = "curFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$Group == "LEX", ]
  tmp$curFuel = factor(tmp$curFuel, levels = c("Smoky","Smokeles","Wood_and_or_Plant"))
  
  form <- as.formula(paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + "), " + curStove"))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, cl_lower, cl_upper, pval, p_val, pval_BHadj)
  
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Smoky (reference)"
  rownames(coefs)[rownames(coefs) == "curFuelSmokeles"] = "Smokeles"
  rownames(coefs)[rownames(coefs) == "curFuelWood_and_or_Plant"] = "Wood_and_or_Plant"

  return(coefs)
}
```

```{r}
sig_color <- c("<= 0.001" = "#7CAE00", "<= 0.01" = "#00BFC4", "<= 0.05" = "#C77CFF", "> 0.05" = "black")

make_forest_plot <- function(df, exposure){
  df$p_val <- as.factor(df$p_val)
  p <- ggplot(data=df, aes(x=EAAs, y=coefficient, ymin=cl_lower, ymax=cl_upper)) +
    geom_point(shape=24, size=2, aes(col = p_val, fill = p_val)) +
    geom_errorbar(shape=24, width=0.3, aes(col = p_val)) +
    geom_hline(yintercept=0, lty=2,color="red") +  # add a dotted line at x=1 after flip
    # ylim(c(-2,2)) +  # not working
    coord_flip() +  # flip coordinates (puts labels on y axis)
    labs(x=expression(log[2]-PFAS))+
    labs(y="Change in year") +
    ggtitle(exposure)+
    theme(plot.title = element_text(size=8.6, hjust = 0.5),
          axis.title.y = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.y = element_text(size=8),
          axis.title.x = element_text(size=8),
          axis.text = element_text(size=7)) +
    scale_color_manual(values = sig_color) +
    scale_fill_manual(values = sig_color)

  return(p)
}

make_forest_plot_lg <- function(df, exposure){
  df$p_val <- as.factor(df$p_val)
  p <- ggplot(data=df, aes(x=EAAs, y=coefficient, ymin=cl_lower, ymax=cl_upper)) +
    geom_point(shape=24, size=2, aes(col = p_val, fill = p_val)) +
    geom_errorbar(shape=24, width=0.3, aes(col = p_val)) +
    geom_hline(yintercept=0, lty=2,color="red") +  # add a dotted line at x=1 after flip
    # ylim(c(-2,2)) +  # not working
    coord_flip() +  # flip coordinates (puts labels on y axis)
    labs(x=expression(log[2]-PFAS))+
    labs(y="Change in year") +
    ggtitle(exposure)+
    theme(plot.title = element_text(size=15, hjust = 0.5),
          axis.title.y = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.y = element_text(size=10),
          axis.title.x = element_text(size=10),
          axis.text = element_text(size=10)) +
    scale_color_manual(values = sig_color) +
    scale_fill_manual(values = sig_color)

  return(p)
}

```

```{r linear_relation_curFuel_mix, fig.width=7, fig.height=6}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 7))
fueltype_smokeless_mean <- data.frame(matrix(NA, ncol = 7))
fueltype_woodplant_mean <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_curFuel_mix(EAAs_vars[i])
    fueltype_smoky_mean[i,] <- rlts[1,]
    fueltype_smokeless_mean[i,] <- rlts[2,]
    fueltype_woodplant_mean[i,] <- rlts[3,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_smokeless_mean) = colnames(rlts)
colnames(fueltype_woodplant_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_smokeless_mean) = EAAs_vars
rownames(fueltype_woodplant_mean) = EAAs_vars

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_woodplant_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky (reference)")
p2 <- make_forest_plot_lg(fueltype_smokeless_mean, "Smokeless")
p3 <- make_forest_plot_lg(fueltype_woodplant_mean, "Wood_and_or_Plant")

plot_list <- list(p1, p2, p3)


for(j in 1:3){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-10, 10))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             layout_matrix = rbind(c(1, NA),
                                   c(2, 3)),
             top = textGrob(paste("Average EAA difference of each current Fuel Type\n compared to Smoky Coal"), gp=gpar(fontsize=15, fontface = 'bold')))

```


## Sensitivity analyses
### Likelihood ratio (LR) test (single model)
Limit the analyses in the primary analysis to include only a single observation from each subject (no need for a mixed model). The rationale for this is that it is not so easy to obtain unbiased p-values from a mixed model for FDR testing. This can be remediated during FDR testing but would be good to check.

Full model: $$Y = \beta_0 + \beta_1 * I(\text{Smokeles}) + \beta_2 * I(\text{Wood_and_or_Plant}) + \epsilon$$
Nested model: $$Y = \beta_0 + \epsilon$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.  
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
preform_lmtest_simple <- function(Y, X, dat = df, log_trans_X = F) {
  tmp <- dat
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")
  
  form_full <- as.formula(paste0(Y," ~ ", X_form))
  lmfit_full <- lm(form_full, data = tmp)
  form_reduced <- as.formula(paste0(Y," ~ 1"))
  lmfit_reduced <- lm(form_reduced, data = tmp)
  
  lrtest <- lrtest(lmfit_full, lmfit_reduced)

  return(round(lrtest$`Pr(>Chisq)`, 4)[2])
}
```

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], X = "curFuel")
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Linear relation
Use a trend test to estimate a linear relation across use categories (1=wood, 2=smokeless coal, 3=smoky coal). 
Fit the equation: $$Y = \beta_0 + \beta_1 * fuel\_type + \epsilon$$
```{r}
preform_regression_curFuel_simple <- function(Y) {
  X = "curFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$Group == "LEX", ]

  tmp$curFuel[tmp$curFuel == "Wood_and_or_Plant"] = 1
  tmp$curFuel[tmp$curFuel == "Smokeles"] = 2
  tmp$curFuel[tmp$curFuel == "Smoky"] = 3
  tmp$curFuel <- as.numeric(tmp$curFuel)

  form <- as.formula(paste0(Y," ~ ", X))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, cl_lower, cl_upper, pval, p_val, pval_BHadj)

  return(coefs)
}
```

```{r linear_relation_curFuel_simple, fig.width=5, fig.height=4}
curFuel_slope <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_curFuel_simple(EAAs_vars[i])
    curFuel_slope[i,] <- rlts[2,]
}

colnames(curFuel_slope) = colnames(rlts)

rownames(curFuel_slope) = EAAs_vars

curFuel_slope$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

print(round(curFuel_slope %>% select(coefficient, std, pval, pval_BHadj), 2) )

make_forest_plot_lg(curFuel_slope, "Slope of current (self-reported) fuel type ") 
ggsave("linear_relation_curFuel_simple.png", dpi=1200)

```








# Cumulative lifetime (self-reported) fuel type 
The numbers of observations with each cumulative lifetime fuel type:
```{r}
table(df$cumFuel)
```

## Primary analysis

### Likelihood ratio (LR) test (mix model)

Full model: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Mix}) \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \beta_6 * curStove + \epsilon
\end{aligned}
$$
Nested model: 
$$
\begin{aligned}
  Y = & \beta_0  \\
  & + \beta_1 * county + \beta_2 * BMI + \beta_3 * ses + \beta_4 * edu + \beta_5 * curStove + \epsilon
\end{aligned}
$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_mix(Y = EAAs_vars[i], X = "cumFuel")
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Linear regression 
In the following section, we performed linear regression with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Mix}) \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \beta_6 * curStove + \epsilon
\end{aligned}
$$  
where $Y$ is one of the epigenetic age accelerations and the reference as the smoky fuel type.  
 
The estimations of $\beta_0$ and $\beta_1$ with given $Y$ are shown below. The $\beta_1$ can be interpreted as "the expected change of Y if switching form the smoky fuel type to the mix fuel type, while holding other variables constant".


```{r}
preform_regression_cumFuel_mix <- function(Y) {
  X = "cumFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$Group == "LEX", ]
  tmp$cumFuel = factor(tmp$cumFuel, levels = c("Smoky","Mix"))
  
  form <- as.formula(paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + "), " + curStove"))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, cl_lower, cl_upper, pval, p_val, pval_BHadj)
  
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Smoky (reference)"
  rownames(coefs)[rownames(coefs) == "cumFuelMix"] = "Mix"

  return(coefs)
}
```


```{r linear_relation_cumFuel_mix, fig.width=7, fig.height=3}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 7))
fueltype_mix_mean <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_cumFuel_mix(EAAs_vars[i])
    fueltype_smoky_mean[i,] <- rlts[1,]
    fueltype_mix_mean[i,] <- rlts[2,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_mix_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_mix_mean) = EAAs_vars

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky (reference)")
p2 <- make_forest_plot_lg(fueltype_mix_mean, "Smokeless")

plot_list <- list(p1, p2)


for(j in 1:2){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-10, 10))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Average EAA difference of each cumulative Fuel Type\n compared to Smoky Coal"), gp=gpar(fontsize=15, fontface = 'bold')))
```


## Sensitivity analyses
### Likelihood ratio (LR) test (single model)

Full model: $$Y = \beta_0 + \beta_1 * I(\text{Mix}) + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], X = "cumFuel")
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Linear relation
Use a trend test to estimate a linear relation across use categories (1=mix, 2=Smoky coal). 
Fit the equation: $$Y = \beta_0 + \beta_1 * fuel\_type + \epsilon$$
```{r}
preform_regression_cumFuel_simple <- function(Y) {
  X = "cumFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$Group == "LEX", ]

  tmp$cumFuel[tmp$cumFuel == "Mix"] = 1
  tmp$cumFuel[tmp$cumFuel == "Smoky"] = 2
  tmp$curFuel <- as.numeric(tmp$curFuel)

  form <- as.formula(paste0(Y," ~ ", X))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, cl_lower, cl_upper, pval, p_val, pval_BHadj)

  return(coefs)
}
```

```{r linear_relation_cumFuel_simple, fig.width=5, fig.height=4}
cumFuel_slope <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_cumFuel_simple(EAAs_vars[i])
    cumFuel_slope[i,] <- rlts[2,]
}

colnames(cumFuel_slope) = colnames(rlts)

rownames(cumFuel_slope) = EAAs_vars

cumFuel_slope$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

print(round(cumFuel_slope %>% select(coefficient, std, pval, pval_BHadj), 2) )

make_forest_plot_lg(cumFuel_slope, "Slope of cumulative (self-reported) fuel type ") 
ggsave("linear_relation_cumFuel_simple.png", dpi=1200)

```






# Childhood  (self-reported) fuel type 
The numbers of observations with each current fuel type:
```{r}
table(df$brthFuel)
```

## Primary analysis

### Likelihood ratio (LR) test (mix model)
Full model: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Wood}) + \beta_2 * I(\text{Smokeles}) + \beta_3 * I(\text{Mix})  \\
  & + \beta_4 * county + \beta_5 * BMI + \beta_6 * ses + \beta_7 * edu + \beta_8 * curStove + \epsilon
\end{aligned}
$$
Nested model: 
$$
\begin{aligned}
  Y = & \beta_0   \\
  & + \beta_1 * county + \beta_2 * BMI + \beta_3 * ses + \beta_4 * edu + \beta_5 * curStove + \epsilon
\end{aligned}
$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_mix(Y = EAAs_vars[i], X = "brthFuel")
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Linear regression 
In the following section, we performed linear regression with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Wood}) + \beta_2 * I(\text{Smokeles}) + \beta_3 * I(\text{Mix})  \\
  & + \beta_4 * county + \beta_5 * BMI + \beta_6 * ses + \beta_7 * edu + \beta_8 * curStove + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and the reference as the smoky fuel type.

The estimations of $\beta_0$, $\beta_1$, $\beta_2$, and $\beta_3$ with given $Y$ are shown below. The $\beta_1$, $\beta_2$, and $\beta_3$ can be interpreted as "the expected change of Y if switching form the smoky fuel type to the given fuel type, while holding other variables constant".


```{r}
preform_regression_brthFuel_mix <- function(Y) {
  X = "brthFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$Group == "LEX", ]
  tmp$brthFuel = factor(tmp$brthFuel, levels = c("Smoky","Wood","Smokeles", "Mix"))
  
  form <- as.formula(paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + "), " + curStove"))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, cl_lower, cl_upper, pval, p_val, pval_BHadj)
  
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Smoky (reference)"
  rownames(coefs)[rownames(coefs) == "brthFuelWood"] = "Wood"
  rownames(coefs)[rownames(coefs) == "brthFuelSmokeles"] = "Smokeles"
  rownames(coefs)[rownames(coefs) == "brthFuelMix"] = "Mix"

  return(coefs)
}
```


```{r linear_relation_brthFuel_mix, fig.width=7, fig.height=6}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 7))
fueltype_wood_mean <- data.frame(matrix(NA, ncol = 7))
fueltype_smokeless_mean <- data.frame(matrix(NA, ncol = 7))
fueltype_mix_mean <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_brthFuel_mix(EAAs_vars[i])
    fueltype_smoky_mean[i,] <- rlts[1,]
    fueltype_wood_mean[i,] <- rlts[2,]
    fueltype_smokeless_mean[i,] <- rlts[3,]
    fueltype_mix_mean[i,] <- rlts[4,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_wood_mean) = colnames(rlts)
colnames(fueltype_smokeless_mean) = colnames(rlts)
colnames(fueltype_mix_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_wood_mean) = EAAs_vars
rownames(fueltype_smokeless_mean) = EAAs_vars
rownames(fueltype_mix_mean) = EAAs_vars

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_wood_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky (reference)")
p2 <- make_forest_plot_lg(fueltype_wood_mean, "Wood")
p3 <- make_forest_plot_lg(fueltype_smokeless_mean, "Smokeless")
p4 <- make_forest_plot_lg(fueltype_mix_mean, "Mix")

plot_list <- list(p1, p2, p3, p4)


for(j in 1:4){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-10, 10))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             layout_matrix = rbind(c(1, 2),
                                   c(3, 4)),
             top = textGrob(paste("Average EAA difference of each Childhood Fuel Type\n compared to Smoky Coal"), gp=gpar(fontsize=15, fontface = 'bold')))

```


## Sensitivity analyses
### Likelihood ratio (LR) test (single model)
Limit the analyses in the primary analysis to include only a single observation from each subject (no need for a mixed model). The rationale for this is that it is not so easy to obtain unbiased p-values from a mixed model for FDR testing. This can be remediated during FDR testing but would be good to check.

Full model: $$Y = \beta_0 + \beta_1 * I(\text{Wood}) + \beta_2 * I(\text{Smokeles}) + \beta_3 * I(\text{Mix}) + \epsilon$$
Nested model: $$Y = \beta_0 + \epsilon$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.    

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], X = "brthFuel")
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Linear relation
Use a trend test to estimate a linear relation across use categories (1=wood, 2=smokeless coal, 3 = mix coal, 4=smoky coal). 
Fit the equation: $$Y = \beta_0 + \beta_1 * fuel\_type + \epsilon$$

```{r}
preform_regression_brthFuel_simple <- function(Y) {
  X = "brthFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$Group == "LEX", ]
  
  tmp$brthFuel[tmp$brthFuel == "Wood"] = 1
  tmp$brthFuel[tmp$brthFuel == "Smokeles"] = 2
  tmp$brthFuel[tmp$brthFuel == "Mix"] = 3
  tmp$brthFuel[tmp$brthFuel == "Smoky"] = 4
  tmp$brthFuel <- as.numeric(tmp$brthFuel)

  form <- as.formula(paste0(Y," ~ ", X))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, cl_lower, cl_upper, pval, p_val, pval_BHadj)

  return(coefs)
}
```

```{r linear_relation_brthFuel_simple, fig.width=5, fig.height=4}
curFuel_slope <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_brthFuel_simple(EAAs_vars[i])
    curFuel_slope[i,] <- rlts[2,]
}

colnames(curFuel_slope) = colnames(rlts)

rownames(curFuel_slope) = EAAs_vars

curFuel_slope$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

print(round(curFuel_slope %>% select(coefficient, std, pval, pval_BHadj), 2) )

make_forest_plot_lg(curFuel_slope, "Slope of cildhood (self-reported) fuel type ") 
ggsave("linear_relation_brthFuel_simple.png", dpi=1200)

```



# Clusters based on model-based exposure estimates at or shortly before the visit (clusCUR6)
The file “LEX_clusCUR6.csv” has information on current pollutant exposures, obtained for the 2 years preceding the visit. To reduce multi-collinearity between exposures, exposure prototypes were derived based on hierarchical cluster analysis in combination followed by principal components analysis. These estimates are available for 6 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:

CUR6_BC_PAH6 – Black carbon (BC) and 6 PAHs   
CUR6_PAH31 – a large cluster of 31 PAHs  
CUR6_NkF – NkF only  
CUR6_PM_RET – Particulate matter (PM) and retene  
CUR6_NO2 – NO2 only  
CUR6_SO2 – SO2 only    



## Summary the exposure estimates:
```{r}
summary(df %>% select(var_name_list$cluster_exp$clusCUR6))
```

## Primary analysis

### Likelihood ratio (LR) test (mix model)

Full model: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{BC_PAH6} + \beta_2 * \text{PAH31} + \beta_3 * \text{NkF} + \beta_4 * \text{PM_RET} + \beta_5 * \text{NO2} + \beta_6 * \text{SO2}\\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_{10} * edu + \beta_{11} * curStove + \epsilon
\end{aligned}
$$
Nested model: 
$$
\begin{aligned}
  Y = & \beta_0  \\
  & + \beta_1 * county + \beta_2 * BMI + \beta_3 * ses + \beta_4 * edu + \beta_5 * curStove + \epsilon
\end{aligned}
$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_mix(Y = EAAs_vars[i], 
                                   X = unname(unlist(var_name_list$cluster_exp$clusCUR6)))
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Linear regression 
In the following section, we performed linear regression with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{BC_PAH6} + \beta_2 * \text{PAH31} + \beta_3 * \text{NkF} + \beta_4 * \text{PM_RET} + \beta_5 * \text{NO2} + \beta_6 * \text{SO2}\\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_{10} * edu + \beta_{11} * curStove + \epsilon
\end{aligned}
$$ 
where $Y$ is one of the epigenetic age accelerations.  
 
The estimations of $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$, and $\beta_6$ with given $Y$ are shown below. The $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$, and $\beta_6$ can be interpreted as "the expected change of Y if increase one unit of given exposure prototype, while holding other variables constant".


```{r}
preform_regression_exposure_prototype <- function(X, Y, log_trans_X=F) {
  
  tmp <- df
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")
  
  tmp <- tmp[tmp$Group == "LEX", ]

  form <- as.formula(paste0(Y," ~ ", X_form, " + ", paste(var_name_list$confounders[-1], collapse  = " + "), " + curStove"))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")
  coefs$pval_BHadj <- p.adjust(coefs$pval, method = "BH")

  coefs <- coefs %>%
    select(coefficient, std, pval, pval_BHadj) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, cl_lower, cl_upper, pval, p_val, pval_BHadj)
  
  return(coefs)
}
```


```{r linear_relation_cumFuel_mix, fig.width=6, fig.height=8}
CUR6_BC_PAH6 <- data.frame(matrix(NA, ncol = 7))
CUR6_PAH31 <- data.frame(matrix(NA, ncol = 7))
CUR6_NkF <- data.frame(matrix(NA, ncol = 7))
CUR6_PM_RET <- data.frame(matrix(NA, ncol = 7))
CUR6_NO2 <- data.frame(matrix(NA, ncol = 7))
CUR6_SO2 <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_exposure_prototype(X = unname(unlist(var_name_list$cluster_exp$clusCUR6)),
                                                 Y = EAAs_vars[i])
    CUR6_BC_PAH6[i,] <- rlts["CUR6_BC_PAH6",]
    CUR6_PAH31[i,] <- rlts["CUR6_PAH31",]
    CUR6_NkF[i,] <- rlts["CUR6_NkF",]
    CUR6_PM_RET[i,] <- rlts["CUR6_PM_RET",]
    CUR6_NO2[i,] <- rlts["CUR6_NO2",]
    CUR6_SO2[i,] <- rlts["CUR6_SO2",]
}

colnames(CUR6_BC_PAH6) = colnames(rlts)
colnames(CUR6_PAH31) = colnames(rlts)
colnames(CUR6_NkF) = colnames(rlts)
colnames(CUR6_PM_RET) = colnames(rlts)
colnames(CUR6_NO2) = colnames(rlts)
colnames(CUR6_SO2) = colnames(rlts)

rownames(CUR6_BC_PAH6) = EAAs_vars
rownames(CUR6_PAH31) = EAAs_vars
rownames(CUR6_NkF) = EAAs_vars
rownames(CUR6_PM_RET) = EAAs_vars
rownames(CUR6_NO2) = EAAs_vars
rownames(CUR6_SO2) = EAAs_vars

CUR6_BC_PAH6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUR6_PAH31$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUR6_NkF$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUR6_PM_RET$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUR6_NO2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUR6_SO2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(CUR6_BC_PAH6, "BC_PAH6")
p2 <- make_forest_plot_lg(CUR6_PAH31, "PAH31")
p3 <- make_forest_plot_lg(CUR6_NkF, "NkF")
p4 <- make_forest_plot_lg(CUR6_PM_RET, "PM_RET")
p5 <- make_forest_plot_lg(CUR6_NO2, "NO2")
p6 <- make_forest_plot_lg(CUR6_SO2, "SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)


for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each exposure prototype \n based on current pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```

## Sensitivity analyses
### Likelihood ratio (LR) test (single model)

Full model: $$Y = \beta_0 + \beta_1 * \text{BC_PAH6} + \beta_2 * \text{PAH31} + \beta_3 * \text{NkF} + \beta_4 * \text{PM_RET} + \beta_5 * \text{NO2} + \beta_6 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCUR6)),
                                      dat = df)
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Likelihood ratio (LR) test (single model) with subjects using only smoky or smokeless coal 

Full model: $$Y = \beta_0 + \beta_1 * \text{BC_PAH6} + \beta_2 * \text{PAH31} + \beta_3 * \text{NkF} + \beta_4 * \text{PM_RET} + \beta_5 * \text{NO2} + \beta_6 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCUR6)),
                                      dat = df[df$curFuel %in% c("Smokeles", "Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Likelihood ratio (LR) test (single model) with subjects only using smoky coal

Full model: $$Y = \beta_0 + \beta_1 * \text{BC_PAH6} + \beta_2 * \text{PAH31} + \beta_3 * \text{NkF} + \beta_4 * \text{PM_RET} + \beta_5 * \text{NO2} + \beta_6 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCUR6)),
                                      dat = df[df$curFuel %in% c("Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```






# Clusters based on model-based exposure estimates accrued before age 18 (clusCHLD5)
The file “LEX_clusCHLD5.csv” has information on estimated pollutant exposures during early childhood. Estimates are available for 5 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:

CHLD5_X7 – a cluster of 7 air pollutants  
CHLD5_X33 – a large cluster of 33 air pollutants  
CHLD5_NkF – NkF only  
CHLD5_NO2 – NO2 only  
CHLD5_SO2 – SO2 only    


## Summary the exposure estimates:
```{r}
summary(df %>% select(var_name_list$cluster_exp$clusCHLD5))
```

## Primary analysis

### Likelihood ratio (LR) test (mix model)

Full model: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{X7} + \beta_2 * \text{X33} + \beta_3 * \text{NkF} + \beta_4 * \text{NO2} + \beta_5 * \text{SO2}\\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_{9} * edu + \beta_{10} * curStove + \epsilon
\end{aligned}
$$
Nested model: 
$$
\begin{aligned}
  Y = & \beta_0  \\
  & + \beta_1 * county + \beta_2 * BMI + \beta_3 * ses + \beta_4 * edu + \beta_5 * curStove + \epsilon
\end{aligned}
$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_mix(Y = EAAs_vars[i], 
                                   X = unname(unlist(var_name_list$cluster_exp$clusCHLD5)))
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Linear regression 
In the following section, we performed linear regression with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{X7} + \beta_2 * \text{X33} + \beta_3 * \text{NkF} + \beta_4 * \text{NO2} + \beta_5 * \text{SO2}\\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_{9} * edu + \beta_{10} * curStove + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.  
 
The estimations of $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, and $\beta_5$ with given $Y$ are shown below. The $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, and $\beta_5$ can be interpreted as "the expected change of Y if increase one unit of given exposure prototype, while holding other variables constant".


```{r linear_relation_cumFuel_mix, fig.width=6, fig.height=8}
CHLD5_X7 <- data.frame(matrix(NA, ncol = 7))
CHLD5_X33 <- data.frame(matrix(NA, ncol = 7))
CHLD5_NkF <- data.frame(matrix(NA, ncol = 7))
CHLD5_NO2 <- data.frame(matrix(NA, ncol = 7))
CHLD5_SO2 <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_exposure_prototype(X = unname(unlist(var_name_list$cluster_exp$clusCHLD5)),
                                                 Y = EAAs_vars[i])
    CHLD5_X7[i,] <- rlts["CHLD5_X7",]
    CHLD5_X33[i,] <- rlts["CHLD5_X33",]
    CHLD5_NkF[i,] <- rlts["CHLD5_NkF",]
    CHLD5_NO2[i,] <- rlts["CHLD5_NO2",]
    CHLD5_SO2[i,] <- rlts["CHLD5_SO2",]
}

colnames(CHLD5_X7) = colnames(rlts)
colnames(CHLD5_X33) = colnames(rlts)
colnames(CHLD5_NkF) = colnames(rlts)
colnames(CHLD5_NO2) = colnames(rlts)
colnames(CHLD5_SO2) = colnames(rlts)

rownames(CHLD5_X7) = EAAs_vars
rownames(CHLD5_X33) = EAAs_vars
rownames(CHLD5_NkF) = EAAs_vars
rownames(CHLD5_NO2) = EAAs_vars
rownames(CHLD5_SO2) = EAAs_vars

CHLD5_X7$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CHLD5_X33$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CHLD5_NkF$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CHLD5_NO2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CHLD5_SO2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(CHLD5_X7, "X7")
p2 <- make_forest_plot_lg(CHLD5_X33, "X33")
p3 <- make_forest_plot_lg(CHLD5_NkF, "NkF")
p4 <- make_forest_plot_lg(CHLD5_NO2, "NO2")
p5 <- make_forest_plot_lg(CHLD5_SO2, "SO2")

plot_list <- list(p1, p2, p3, p4, p5)


for(j in 1:5){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each exposure prototype \n based on childhood pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```

## Sensitivity analyses
### Likelihood ratio (LR) test (single model)

Full model: $$Y = \beta_0 + \beta_1 * \text{X7} + \beta_2 * \text{X33} + \beta_3 * \text{NkF} + \beta_4 * \text{NO2} + \beta_5 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCHLD5)),
                                      dat = df)
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Likelihood ratio (LR) test (single model) with subjects using only smoky or smokeless coal 

Full model: $$Y = \beta_0 + \beta_1 * \text{X7} + \beta_2 * \text{X33} + \beta_3 * \text{NkF} + \beta_4 * \text{NO2} + \beta_5 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCHLD5)),
                                      dat = df[df$curFuel %in% c("Smokeles", "Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Likelihood ratio (LR) test (single model) with subjects only using smoky coal

Full model: $$Y = \beta_0 + \beta_1 * \text{X7} + \beta_2 * \text{X33} + \beta_3 * \text{NkF} + \beta_4 * \text{NO2} + \beta_5 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCHLD5)),
                                      dat = df[df$curFuel %in% c("Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```







# Clusters based on model-based lifetime exposure estimates (clusCUM6)
The file “LEX_clus CUM6.csv” has information on estimated cumulative pollutant exposures during the lifecourse. Estimates are available for 6 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:  

CUM6_BC_NO2_PM – a cluster of BC, NO2, and PM  
CUM6_PAH36 – a large cluster of 36 PAHs  
CUM6_DlP – DlP only  
CUM6_NkF – NkF only  
CUM6_RET – retene only  
CUM6_SO2 – SO2 only   
 

## Summary the exposure estimates:
```{r}
summary(df %>% select(var_name_list$cluster_exp$clusCUM6))
```

## Primary analysis

### Likelihood ratio (LR) test (mix model)

Full model: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{BC_NO2_PM} + \beta_2 * \text{PAH36} + \beta_3 * \text{DlP} + \beta_4 * \text{NkF} + \beta_5 * \text{RET} + \beta_6 * \text{SO2}\\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_{10} * edu + \beta_{11} * curStove + \epsilon
\end{aligned}
$$
Nested model: 
$$
\begin{aligned}
  Y = & \beta_0  \\
  & + \beta_1 * county + \beta_2 * BMI + \beta_3 * ses + \beta_4 * edu + \beta_5 * curStove + \epsilon
\end{aligned}
$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_mix(Y = EAAs_vars[i], 
                                   X = unname(unlist(var_name_list$cluster_exp$clusCUM6)))
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Linear regression 
In the following section, we performed linear regression with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{BC_NO2_PM} + \beta_2 * \text{PAH36} + \beta_3 * \text{DlP} + \beta_4 * \text{NkF} + \beta_5 * \text{RET} + \beta_6 * \text{SO2}\\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_{10} * edu + \beta_{11} * curStove + \epsilon
\end{aligned}
$$ 
where $Y$ is one of the epigenetic age accelerations.  
 
The estimations of $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$, and $\beta_6$ with given $Y$ are shown below. The $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$, and $\beta_6$ can be interpreted as "the expected change of Y if increase one unit of given exposure prototype, while holding other variables constant".


```{r linear_relation_cumFuel_mix, fig.width=6, fig.height=8}
CUM6_BC_NO2_PM <- data.frame(matrix(NA, ncol = 7))
CUM6_PAH36 <- data.frame(matrix(NA, ncol = 7))
CUM6_DlP <- data.frame(matrix(NA, ncol = 7))
CUM6_NkF <- data.frame(matrix(NA, ncol = 7))
CUM6_RET <- data.frame(matrix(NA, ncol = 7))
CUM6_SO2 <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_exposure_prototype(X = unname(unlist(var_name_list$cluster_exp$clusCUM6)),
                                                 Y = EAAs_vars[i])
    CUM6_BC_NO2_PM[i,] <- rlts["CUM6_BC_NO2_PM",]
    CUM6_PAH36[i,] <- rlts["CUM6_PAH36",]
    CUM6_DlP[i,] <- rlts["CUM6_DlP",]
    CUM6_NkF[i,] <- rlts["CUM6_NkF",]
    CUM6_RET[i,] <- rlts["CUM6_RET",]
    CUM6_SO2[i,] <- rlts["CUM6_SO2",]
}

colnames(CUM6_BC_NO2_PM) = colnames(rlts)
colnames(CUM6_PAH36) = colnames(rlts)
colnames(CUM6_DlP) = colnames(rlts)
colnames(CUM6_NkF) = colnames(rlts)
colnames(CUM6_RET) = colnames(rlts)
colnames(CUM6_SO2) = colnames(rlts)

rownames(CUM6_BC_NO2_PM) = EAAs_vars
rownames(CUM6_PAH36) = EAAs_vars
rownames(CUM6_DlP) = EAAs_vars
rownames(CUM6_NkF) = EAAs_vars
rownames(CUM6_RET) = EAAs_vars
rownames(CUM6_SO2) = EAAs_vars

CUM6_BC_NO2_PM$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUM6_PAH36$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUM6_DlP$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUM6_NkF$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUM6_RET$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
CUM6_SO2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(CUM6_BC_NO2_PM, "BC_NO2_PM")
p2 <- make_forest_plot_lg(CUM6_PAH36, "PAH36")
p3 <- make_forest_plot_lg(CUM6_DlP, "DlP")
p4 <- make_forest_plot_lg(CUM6_NkF, "NkF")
p5 <- make_forest_plot_lg(CUM6_RET, "RET")
p6 <- make_forest_plot_lg(CUM6_SO2, "SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)


for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each exposure prototype \n based on lifetime pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```

## Sensitivity analyses
### Likelihood ratio (LR) test (single model)

Full model: $$Y = \beta_0 + \beta_1 * \text{BC_NO2_PM} + \beta_2 * \text{PAH36} + \beta_3 * \text{DlP} + \beta_4 * \text{NkF} + \beta_5 * \text{RET} + \beta_6 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCUM6)),
                                      dat = df)
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Likelihood ratio (LR) test (single model) with subjects using only smoky or smokeless coal 

Full model: $$Y = \beta_0 + \beta_1 * \text{BC_NO2_PM} + \beta_2 * \text{PAH36} + \beta_3 * \text{DlP} + \beta_4 * \text{NkF} + \beta_5 * \text{RET} + \beta_6 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$   

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCUM6)),
                                      dat = df[df$curFuel %in% c("Smokeles", "Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Likelihood ratio (LR) test (single model) with subjects only using smoky coal

Full model: $$Y = \beta_0 + \beta_1 * \text{BC_NO2_PM} + \beta_2 * \text{PAH36} + \beta_3 * \text{DlP} + \beta_4 * \text{NkF} + \beta_5 * \text{RET} + \beta_6 * \text{SO2} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusCUM6)),
                                      dat = df[df$curFuel %in% c("Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```








# Clusters based on pollutant measurements (clusMEAS6)
# Clusters based on urinary biomarkers (clusURI5)
The file “LEX_clusURI5.csv” has information on measured urinary biomarkers obtained during each visit. Estimates are available for 5 different prototypes (cluster variables) for a total of 163 subjects and 186 visits. The prototypes are labelled as:  


URI5_NAP_1M_2M – a cluster of Naphthalene, 1Methylnaphthalene, and 2Methylnaphthalene  
URI5_ACE – Acenaphthene only  
URI5_FLU_PHE – Fluoranthene and Phenanthrene_anth  
URI5_PYR – Pyrene only  
URI5_CHR – Baa_Chrysene only  
 



## Summary the exposure estimates:
```{r}
summary(df %>% select(var_name_list$cluster_exp$clusURI5))
```

## Primary analysis

### Likelihood ratio (LR) test (mix model)

Full model: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{NAP_1M_2M} + \beta_2 * \text{ACE} + \beta_3 * \text{FLU_PHE} + \beta_4 * \text{PYR} + \beta_5 * \text{CHR}\\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_{9} * edu + \beta_{10} * curStove + \epsilon
\end{aligned}
$$
Nested model: 
$$
\begin{aligned}
  Y = & \beta_0  \\
  & + \beta_1 * county + \beta_2 * BMI + \beta_3 * ses + \beta_4 * edu + \beta_5 * curStove + \epsilon
\end{aligned}
$$
$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.  

P-values results: 

```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_mix(Y = EAAs_vars[i], 
                                   X = unname(unlist(var_name_list$cluster_exp$clusURI5)))
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Linear regression 
In the following section, we performed linear regression with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * \text{NAP_1M_2M} + \beta_2 * \text{ACE} + \beta_3 * \text{FLU_PHE} + \beta_4 * \text{PYR} + \beta_5 * \text{CHR}\\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_{9} * edu + \beta_{10} * curStove + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.  
 
The estimations of $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, and $\beta_5$ with given $Y$ are shown below. The $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, and $\beta_5$ can be interpreted as "the expected change of Y if increase one unit of given exposure prototype, while holding other variables constant".


```{r linear_relation_cumFuel_mix, fig.width=6, fig.height=8}
URI5_NAP_1M_2M <- data.frame(matrix(NA, ncol = 7))
URI5_ACE <- data.frame(matrix(NA, ncol = 7))
URI5_FLU_PHE <- data.frame(matrix(NA, ncol = 7))
URI5_PYR <- data.frame(matrix(NA, ncol = 7))
URI5_CHR <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_exposure_prototype(X = unname(unlist(var_name_list$cluster_exp$clusURI5)),
                                                 Y = EAAs_vars[i])
    URI5_NAP_1M_2M[i,] <- rlts["URI5_NAP_1M_2M",]
    URI5_ACE[i,] <- rlts["URI5_ACE",]
    URI5_FLU_PHE[i,] <- rlts["URI5_FLU_PHE",]
    URI5_PYR[i,] <- rlts["URI5_PYR",]
    URI5_CHR[i,] <- rlts["URI5_CHR",]
}

colnames(URI5_NAP_1M_2M) = colnames(rlts)
colnames(URI5_ACE) = colnames(rlts)
colnames(URI5_FLU_PHE) = colnames(rlts)
colnames(URI5_PYR) = colnames(rlts)
colnames(URI5_CHR) = colnames(rlts)

rownames(URI5_NAP_1M_2M) = EAAs_vars
rownames(URI5_ACE) = EAAs_vars
rownames(URI5_FLU_PHE) = EAAs_vars
rownames(URI5_PYR) = EAAs_vars
rownames(URI5_CHR) = EAAs_vars

URI5_NAP_1M_2M$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
URI5_ACE$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
URI5_FLU_PHE$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
URI5_PYR$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
URI5_CHR$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(URI5_NAP_1M_2M, "X7")
p2 <- make_forest_plot_lg(URI5_ACE, "X33")
p3 <- make_forest_plot_lg(URI5_FLU_PHE, "NkF")
p4 <- make_forest_plot_lg(URI5_PYR, "NO2")
p5 <- make_forest_plot_lg(URI5_CHR, "SO2")

plot_list <- list(p1, p2, p3, p4, p5)


for(j in 1:5){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each exposure prototype \n based on urinary biomarkers"), gp=gpar(fontsize=15, fontface = 'bold')))
```

## Sensitivity analyses
### Likelihood ratio (LR) test (single model)

Full model: $$Y = \beta_0 + \beta_1 * \text{NAP_1M_2M} + \beta_2 * \text{ACE} + \beta_3 * \text{FLU_PHE} + \beta_4 * \text{PYR} + \beta_5 * \text{CHR} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusURI5)),
                                      dat = df)
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```

### Likelihood ratio (LR) test (single model) with subjects using only smoky or smokeless coal 

Full model: $$Y = \beta_0 + \beta_1 * \text{NAP_1M_2M} + \beta_2 * \text{ACE} + \beta_3 * \text{FLU_PHE} + \beta_4 * \text{PYR} + \beta_5 * \text{CHR} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusURI5)),
                                      dat = df[df$curFuel %in% c("Smokeles", "Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```


### Likelihood ratio (LR) test (single model) with subjects only using smoky coal

Full model: $$Y = \beta_0 + \beta_1 * \text{NAP_1M_2M} + \beta_2 * \text{ACE} + \beta_3 * \text{FLU_PHE} + \beta_4 * \text{PYR} + \beta_5 * \text{CHR} + \epsilon$$  
Nested model: $$Y = \beta_0 + \epsilon$$  

$H_0$: The full model and the nested model fit the data equally well. Thus, you should use the nested model.   
$H_A$: The full model fits the data significantly better than the nested model. Thus, you should use the full model.

P-values results: 
```{r}
p_vals <- rep(NA, length(EAAs_vars))
for(i in 1:length(EAAs_vars)){
    p_vals[i] = preform_lmtest_simple(Y = EAAs_vars[i], 
                                      X = unname(unlist(var_name_list$cluster_exp$clusURI5)),
                                      dat = df[df$curFuel %in% c("Smoky"), ])
}
p_vals_BHadj <- p.adjust(p_vals, method = "BH")
p_vals_mat <- cbind(p_vals, p_vals_BHadj)
rownames(p_vals_mat) <- EAAs_names

print(round(p_vals_mat,4))
```









# Ambient Exposure
## Linear regression (simple model)

In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 *X + \epsilon$$
where $Y$ is one of the epigenetic age accelerations, and $X$ is one of the ambient exposure measurements.

The estimations of $\beta_1$ with given $Y$ and $X$ are shown below, which can be interpreted as "the mean of Y changes given a one-unit increase in X while holding other variables constant".
```{r}
preform_regression <- function(Y, X, log_trans_X = F) {
  tmp <- df[!is.na(df[,X]),]
  if(log_trans_X) {
    tmp[,X] <- log(tmp[,X])
  }
  form <- as.formula(paste0(Y,"~",X))
  lmfit <- lm(form, data = tmp)
  coef <- round(summary(lmfit)$coefficients[2,1],4)
  coef_text <- as.character(coef)
  std <- round(summary(lmfit)$coefficients[2,2],4)
  std_text <- as.character(std)
  sig_star <- ifelse(summary(lmfit)$coefficients[2,4]<0.001, "***",
                ifelse(summary(lmfit)$coefficients[2,4]<0.01, "** ",
                       ifelse(summary(lmfit)$coefficients[2,4]<0.05, "*  ", "   ")))
  sig <- ifelse(summary(lmfit)$coefficients[2,4] <= 0.001, "<= 0.001",
                ifelse(summary(lmfit)$coefficients[2,4] <= 0.01, "<= 0.01",
                       ifelse(summary(lmfit)$coefficients[2,4] <= 0.05, "<= 0.05", "> 0.05")))
  rlt <- paste0(coef_text, paste0(rep(" ", 7-nchar(coef_text)), collapse = ""),
                sig_star,
                "(", std_text, paste0(rep(" ", 6-nchar(std_text)), collapse = ""), ")")
  cl_l <- coef - 1.96 * std
  cl_u <- coef + 1.96 * std
  return(c(rlt, coef, cl_l, cl_u, sig))
}


options(scipen = 999)
```

```{r linear_reg_coef_ambient, fig.width=10, fig.height=16}
lm_coef_ambient <- data.frame()
lm_coef_ambient_plots <- list()

for(j in 1:length(ambient_exp_vars)){
  plot_df <- data.frame(matrix(NA, ncol = 4))
  for(i in 1:length(EAAs_vars)){
    rlts = preform_regression(Y = EAAs_vars[i], X = ambient_exp_vars[j])
    lm_coef_ambient[i,j] <- rlts[1]
    plot_df[i,] <- rlts[c(2,3,4,5)]
  }
  colnames(plot_df) <- c("coefficient", "cl_lower", "cl_upper", "p_val")
  rownames(plot_df) <- EAAs_vars
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$cl_lower = as.numeric(plot_df$cl_lower)
  plot_df$cl_upper = as.numeric(plot_df$cl_upper)

  pj <- make_forest_plot(plot_df, ambient_exp_label[j])
  lm_coef_ambient_plots[[j]] = pj
}


lm_coef_ambient[9,1] <- c("[P<0.001: ***; P<0.01: **; P<0.05: *]")
colnames(lm_coef_ambient) <- ambient_exp_names
rownames(lm_coef_ambient) <- c(EAAs_names, "")


grid.arrange(grobs = lm_coef_ambient_plots,  ncol = 3,
             top = textGrob(paste("Ambient Exposures"), gp=gpar(fontsize=15, fontface = 'bold')))

for(j in 1:length(ambient_exp_vars)){
  lm_coef_ambient_plots[[j]] <- lm_coef_ambient_plots[[j]] + ylim(c(-0.055, 0.055))
}
grid.arrange(grobs = lm_coef_ambient_plots,  ncol = 3,
             top = textGrob(paste("Ambient Exposures (same range)"), gp=gpar(fontsize=15, fontface = 'bold')))

cat("X: Ambient Exposure Measurements:")
print(lm_coef_ambient)
```

```{r}
write.csv(lm_coef_ambient, "../data/2.analysis/lm_coefficients_ambient.csv")
```

# Urinary Measurements
## Linear regression (simple model)
In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 *X + \epsilon$$
where $Y$ is one of the epigenetic age accelerations, and $X$ is one of the urinary measurements.

The estimations of $\beta_1$ with given $Y$ and $X$ are shown below, which can be interpreted as "the mean of Y changes given a one-unit increase in X while holding other variables constant".
```{r linear_reg_coef_urine, fig.width=10, fig.height=9}
lm_coef_urine <- data.frame()
lm_coef_urine_plots <- list()

for(j in 1:length(urinary_exp_vars)){
  plot_df <- data.frame(matrix(NA, ncol = 4))
  for(i in 1:length(EAAs_vars)){
    rlts = preform_regression(Y = EAAs_vars[i], X = urinary_exp_vars[j])
    lm_coef_urine[i,j] <- rlts[1]
    plot_df[i,] <- rlts[c(2,3,4,5)]
  }
  colnames(plot_df) <- c("coefficient", "cl_lower", "cl_upper", "p_val")
  rownames(plot_df) <- EAAs_vars
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$cl_lower = as.numeric(plot_df$cl_lower)
  plot_df$cl_upper = as.numeric(plot_df$cl_upper)

  pj <- make_forest_plot(plot_df, urinary_exp_label[j])
  lm_coef_urine_plots[[j]] = pj
}


lm_coef_urine[9,1] <- c("[P<0.001: ***; P<0.01: **; P<0.05: *]")
colnames(lm_coef_urine) <- urinary_exp_names
rownames(lm_coef_urine) <- c(EAAs_names, "")


grid.arrange(grobs = lm_coef_urine_plots,  ncol = 3,
             top = textGrob(paste("Urinary Measurements"), gp=gpar(fontsize=15, fontface = 'bold')))

for(j in 1:length(urinary_exp_vars)){
  lm_coef_urine_plots[[j]] <- lm_coef_urine_plots[[j]] + ylim(c(-1.5, 3))
}
grid.arrange(grobs = lm_coef_urine_plots,  ncol = 3,
             top = textGrob(paste("Urinary Measurements (same range)"), gp=gpar(fontsize=15, fontface = 'bold')))

cat("X: Urinary Measurements:")
print(lm_coef_urine)

```


```{r}
write.csv(lm_coef_urine, "../data/2.analysis/lm_coefficients_urine.csv")
```






















```{r}
knitr::knit_exit()
```


# Not knitted: 

# Linear regression
In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 *X$$
where $Y$ is one of the epigenetic age accelerations, and $X$ is one of the ambient exposure measurements or the urinary measurements.

The estimations of $\beta_1$ with given $Y$ and $X$ are shown below, which can be interpreted as "the mean of Y changes given a one-unit increase in X while holding other variables constant".



## Current fuel type
In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 * I(\text{Smokeles}) + \beta_2 * I(\text{Smoky}) + \beta_3 * I(\text{Wood_and_or_Plant})$$
where $Y$ is one of the epigenetic age accelerations and the reference as the gas/electricity fuel type from the control group.

The estimations of $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$ with given $Y$ are shown below, which can be interpreted as "the expected change of Y if switching form the gas/electricity fuel type to the given fuel type, while holding other variables constant".

The numbers of observations with each current fuel type:
```{r}
table(df$curFuel)
```

```{r}
preform_regression_curFuel <- function(Y) {
  X = "curFuel"
  tmp <- df[!is.na(df[,X]),]

  tmp <- tmp[tmp$Group == "LEX", ]
  # tmp$fueltype = factor(tmp$fueltype, levels = c("Gas/Elect","Smokeles","Smoky","Wood_and_or_Plant"))
  tmp$fueltype = factor(tmp$fueltype, levels = c("Smoky","Smokeles","Wood_and_or_Plant"))

  form <- as.formula(paste0(Y,"~",X, "+", "curStove + county + BMI + ses + edu"))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")

  coefs <- coefs %>%
    select(coefficient, std, pval) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05"))))
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Gas/Elect"
  rownames(coefs)[rownames(coefs) == "curFuelSmokeles"] = "Smokeles"
  rownames(coefs)[rownames(coefs) == "curFuelSmoky"] = "Smoky"
  rownames(coefs)[rownames(coefs) == "curFuelWood_and_or_Plant"] = "Wood_and_or_Plant"

  return(coefs)
}
```

```{r linear_regression_fueltype_mean, fig.width=7, fig.height=5}
fueltype_gas_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smokeless_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_woodplant_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_curFuel(EAAs_vars[i])
    fueltype_gas_mean[i,] <- rlts[1,]
    fueltype_smokeless_mean[i,] <- rlts[2,]
    fueltype_smoky_mean[i,] <- rlts[3,]
    fueltype_woodplant_mean[i,] <- rlts[4,]
}

colnames(fueltype_gas_mean) = colnames(rlts)
colnames(fueltype_smokeless_mean) = colnames(rlts)
colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_woodplant_mean) = colnames(rlts)

rownames(fueltype_gas_mean) = EAAs_vars
rownames(fueltype_smokeless_mean) = EAAs_vars
rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_woodplant_mean) = EAAs_vars

fueltype_gas_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_woodplant_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot(fueltype_gas_mean, "Gas/Electricity (reference)")
p2 <- make_forest_plot(fueltype_smokeless_mean, "Smokeless")
p3 <- make_forest_plot(fueltype_smoky_mean, "Smoky")
p4 <- make_forest_plot(fueltype_woodplant_mean, "Wood_and_or_Plant")

plot_list <- list(p1, p2, p3, p4)


for(j in 1:4){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-12, 12))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Average EAA difference of each current Fuel Type\n compared to Gas/Electricity"), gp=gpar(fontsize=15, fontface = 'bold')))

```

## Childhood Exposure: Birth Fuel Type
In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 * I(Smokeles) + \beta_2 * I(Smoky) + \beta_3 * I(Wood) + \beta_4 * I(Mix)$$
where $Y$ is one of the epigenetic age accelerations and the reference as the gas/electricity fuel type from the control group.

The estimations of $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ with given $Y$ are shown below, which can be interpreted as "the expected change of Y if switching form the gas/electricity fuel type to the given fuel type, while holding other variables constant".


The numbers of observations with each current fuel type:
```{r}
table(df$brthFuel)
```

```{r}
preform_regression_brthFuel <- function(Y) {
  X = "brthFuel"
  tmp <- df[!is.na(df[,X]),]

  tmp$brthFuel = factor(tmp$brthFuel, levels = c("Gas/Elect", "Smokeles","Smoky","Wood","Mix"))

  form <- as.formula(paste0(Y,"~",X))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")

  coefs <- coefs %>%
    select(coefficient, std, pval) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05"))))
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Gas/Elect"
  rownames(coefs)[rownames(coefs) == "brthFuelSmokeles"] = "Smokeles"
  rownames(coefs)[rownames(coefs) == "brthFuelSmoky"] = "Smoky"
  rownames(coefs)[rownames(coefs) == "brthFuelWood"] = "Wood"
  rownames(coefs)[rownames(coefs) == "brthFuelMix"] = "Mix"

  return(coefs)
}
```

```{r linear_regression_birth_fueltype_mean, fig.width=7, fig.height=7.5}
fueltype_gas_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smokeless_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_wood_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_mix_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_brthFuel(EAAs_vars[i])
    fueltype_gas_mean[i,] <- rlts[1,]
    fueltype_smokeless_mean[i,] <- rlts[2,]
    fueltype_smoky_mean[i,] <- rlts[3,]
    fueltype_wood_mean[i,] <- rlts[4,]
    fueltype_mix_mean[i,] <- rlts[5,]
}

colnames(fueltype_gas_mean) = colnames(rlts)
colnames(fueltype_mix_mean) = colnames(rlts)
colnames(fueltype_smokeless_mean) = colnames(rlts)
colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_wood_mean) = colnames(rlts)

rownames(fueltype_gas_mean) = EAAs_vars
rownames(fueltype_mix_mean) = EAAs_vars
rownames(fueltype_smokeless_mean) = EAAs_vars
rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_wood_mean) = EAAs_vars

fueltype_gas_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_wood_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p_gas <- make_forest_plot(fueltype_gas_mean, "Gas/Electricity (reference)")
p_smokeles <- make_forest_plot(fueltype_smokeless_mean, "Smokeless")
p_smoky <- make_forest_plot(fueltype_smoky_mean, "Smoky")
p_wood <- make_forest_plot(fueltype_wood_mean, "Wood")
p_mix <- make_forest_plot(fueltype_mix_mean, "Mix")


plot_list <- list(p_gas, p_smokeles, p_smoky, p_wood, p_mix)


for(j in 1:5){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-12, 12))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             layout_matrix = rbind(c(1, NA),
                                   c(2,3),
                                   c(4,5)),
             top = textGrob(paste("Average EAA difference of each Birth Fuel Type \ncompared to Gas/Electricity"), gp=gpar(fontsize=13, fontface = 'bold')))

```

## Lifetime Exposure: Cumulative Fuel Type
In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 * I(Smoky) + \beta_2 * I(Mix)$$
where $Y$ is one of the epigenetic age accelerations and the reference as the gas/electricity fuel type from the control group.

The estimations of $\beta_0$, $\beta_1$, and $\beta_2$ with given $Y$ are shown below, which can be interpreted as "the expected change of Y if switching form the gas/electricity fuel type to the given fuel type, while holding other variables constant".


The numbers of observations with each cumulative fuel type:
```{r}
table(df$cumFuel)
```


```{r}
preform_regression_cumFuel <- function(Y) {
  X = "cumFuel"
  tmp <- df[!is.na(df[,X]),]

  tmp$cumFuel = factor(tmp$cumFuel, levels = c("Gas/Elect", "Smoky", "Mix"))

  form <- as.formula(paste0(Y,"~",X))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")

  coefs <- coefs %>%
    select(coefficient, std, pval) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05"))))
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Gas/Elect"
  rownames(coefs)[rownames(coefs) == "cumFuelSmoky"] = "Smoky"
  rownames(coefs)[rownames(coefs) == "cumFuelMix"] = "Mix"

  return(coefs)
}
```

```{r linear_regression_lifetime_fueltype_mean, fig.width=10.5, fig.height=5}
fueltype_gas_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_mix_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_cumFuel(EAAs_vars[i])
    fueltype_gas_mean[i,] <- rlts[1,]
    fueltype_smoky_mean[i,] <- rlts[2,]
    fueltype_mix_mean[i,] <- rlts[3,]
}

colnames(fueltype_gas_mean) = colnames(rlts)
colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_mix_mean) = colnames(rlts)

rownames(fueltype_gas_mean) = EAAs_vars
rownames(fueltype_mix_mean) = EAAs_vars
rownames(fueltype_smoky_mean) = EAAs_vars

fueltype_gas_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p_gas <- make_forest_plot(fueltype_gas_mean, "Gas/Electricity (reference)")
p_smoky <- make_forest_plot(fueltype_smoky_mean, "Smoky")
p_mix <- make_forest_plot(fueltype_wood_mean, "Mix")

plot_list <- list(p_gas, p_smoky, p_mix)


for(j in 1:3){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-12, 12))
}
grid.arrange(grobs = plot_list,  ncol = 3,
             top = textGrob(paste("Average EAA difference of each Lifetime Fuel Type \ncompared to Gas/Electricity"), gp=gpar(fontsize=15, fontface = 'bold')))

```



## Current Stove Type
```{r}
table(df$curStove)
```
In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 * I(\text{Portable_stove}) + \beta_2 * I(\text{Ventilated}) + \beta_3 * I(\text{Firepit_and_unventilated}) + \beta_4 * I(\text{Mix})$$
where $Y$ is one of the epigenetic age accelerations and the baseline as the Gas/Elect stove type from the control group.

The estimations of $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ with given $Y$ are shown below, which can be interpreted as "the expected change of Y if switching form the Gas/Electricity stove type to the given stove type, while holding other variables constant".

```{r}
preform_regression_curStove <- function(Y) {
  X = "curStove"
  tmp <- df[!is.na(df[,X]),]

  tmp$curStove = factor(tmp$curStove, levels = c("Gas/Elect", "Portable_stove", "Ventilated", "Firepit_and_unventilated", "Mix"))

  form <- as.formula(paste0(Y,"~",X))
  lmfit <- lm(form, data = tmp)

  coefs <- as.data.frame(summary(lmfit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")

  coefs <- coefs %>%
    select(coefficient, std, pval) %>%
    mutate(cl_lower = coefficient - 1.96 * std,
           cl_upper = coefficient + 1.96 * std,
           p_val = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05"))))
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Gas/Elect"
  rownames(coefs)[rownames(coefs) == "curStovePortable_stove"] = "Portable_stove"
  rownames(coefs)[rownames(coefs) == "curStoveVentilated"] = "Ventilated"
  rownames(coefs)[rownames(coefs) == "curStoveFirepit_and_unventilated"] = "Firepit_and_unventilated"
  rownames(coefs)[rownames(coefs) == "curStoveMix"] = "Mix"

  return(coefs)
}
```

```{r linear_regression_stovetype_mean, fig.width=7, fig.height=7.5}
stovetype_gas_mean <- data.frame(matrix(NA, ncol = 6))
stovetype_port_mean <- data.frame(matrix(NA, ncol = 6))
stovetype_vent_mean <- data.frame(matrix(NA, ncol = 6))
stovetype_fire_mean <- data.frame(matrix(NA, ncol = 6))
stovetype_mix_mean <- data.frame(matrix(NA, ncol = 6))


for(i in 1:length(EAAs_vars)){
    rlts = preform_regression_curStove(EAAs_vars[i])
    stovetype_gas_mean[i,] <- rlts[1,]
    stovetype_port_mean[i,] <- rlts[2,]
    stovetype_vent_mean[i,] <- rlts[3,]
    stovetype_fire_mean[i,] <- rlts[4,]
    stovetype_mix_mean[i,] <- rlts[5,]
}

colnames(stovetype_gas_mean) = colnames(rlts)
colnames(stovetype_port_mean) = colnames(rlts)
colnames(stovetype_vent_mean) = colnames(rlts)
colnames(stovetype_fire_mean) = colnames(rlts)
colnames(stovetype_mix_mean) = colnames(rlts)

rownames(stovetype_gas_mean) = EAAs_vars
rownames(stovetype_port_mean) = EAAs_vars
rownames(stovetype_vent_mean) = EAAs_vars
rownames(stovetype_fire_mean) = EAAs_vars
rownames(stovetype_mix_mean) = EAAs_vars

stovetype_gas_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
stovetype_port_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
stovetype_vent_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
stovetype_fire_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
stovetype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot(stovetype_gas_mean, "Gas/Electricity (reference)")
p2 <- make_forest_plot(stovetype_port_mean, "Portable_stove")
p3 <- make_forest_plot(stovetype_vent_mean, "Ventilated")
p4 <- make_forest_plot(stovetype_fire_mean, "Firepit_and_unventilated")
p5 <- make_forest_plot(stovetype_mix_mean, "Mix")

plot_list <- list(p1, p2, p3, p4, p5)

# grid.arrange(grobs = plot_list,  ncol = 2,
#              top = textGrob(paste("Average EAA of each Fuel Type"), gp=gpar(fontsize=15, fontface = 'bold')))

for(j in 1:5){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-12, 12))
}
grid.arrange(grobs = plot_list,  ncol = 2,
             layout_matrix = rbind(c(1, NA),
                                   c(2,3),
                                   c(4,5)),
             top = textGrob(paste("Average EAA difference of each Stove Type \ncompared to the control group"), gp=gpar(fontsize=15, fontface = 'bold')))

```



<!-- ## Stove type -->
<!-- In the following section, we performed linear regression with equation $$Y = \beta_0 + \beta_1 * I(HS) + \beta_2 * I(LS) + \beta_3 * I(Mix) + \beta_4 * I(PS)$$ -->
<!-- where $Y$ is one of the epigenetic age accelerations and the baseline as the FP stove type. -->

<!-- The estimations of $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ with given $Y$ are shown below, which can be interpreted as "the expected change of Y if switching form the FP stove type to the given stove type, while holding other variables constant". -->

<!-- ```{r} -->
<!-- table(df$stovetype) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- preform_regression_fueltype <- function(Y) { -->
<!--   X = "stovetype" -->
<!--   tmp <- df[!is.na(df[,X]),] -->
<!--   form <- as.formula(paste0(Y,"~",X)) -->
<!--   lmfit <- lm(form, data = tmp) -->

<!--   coefs <- as.data.frame(summary(lmfit)$coefficients) -->
<!--   colnames(coefs) <- c("coefficient", "std", "t_val", "pval") -->

<!--   coefs <- coefs %>% -->
<!--     select(coefficient, std, pval) %>% -->
<!--     mutate(cl_lower = coefficient - 1.96 * std, -->
<!--            cl_upper = coefficient + 1.96 * std, -->
<!--            p_val = ifelse(pval <= 0.001, "<= 0.001", -->
<!--                         ifelse(pval <= 0.01, "<= 0.01", -->
<!--                                ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) -->
<!--   rownames(coefs)[rownames(coefs) == "(Intercept)"] = "FP" -->
<!--   rownames(coefs)[rownames(coefs) == "stovetypeHS"] = "HS" -->
<!--   rownames(coefs)[rownames(coefs) == "stovetypeLS"] = "LS" -->
<!--   rownames(coefs)[rownames(coefs) == "stovetypeMix"] = "Mix" -->
<!--   rownames(coefs)[rownames(coefs) == "stovetypePS"] = "PS" -->

<!--   return(coefs) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r linear_regression_stovetype_mean, fig.width=7, fig.height=7.5} -->
<!-- stovetype_FP_mean <- data.frame(matrix(NA, ncol = 6)) -->
<!-- stovetype_HS_mean <- data.frame(matrix(NA, ncol = 6)) -->
<!-- stovetype_LS_mean <- data.frame(matrix(NA, ncol = 6)) -->
<!-- stovetype_Mix_mean <- data.frame(matrix(NA, ncol = 6)) -->
<!-- stovetype_PS_mean <- data.frame(matrix(NA, ncol = 6)) -->


<!-- for(i in 1:length(EAAs_vars)){ -->
<!--     rlts = preform_regression_fueltype(EAAs_vars[i]) -->
<!--     stovetype_FP_mean[i,] <- rlts[1,] -->
<!--     stovetype_HS_mean[i,] <- rlts[2,] -->
<!--     stovetype_LS_mean[i,] <- rlts[3,] -->
<!--     stovetype_Mix_mean[i,] <- rlts[4,] -->
<!--     stovetype_PS_mean[i,] <- rlts[5,] -->
<!-- } -->

<!-- colnames(stovetype_FP_mean) = colnames(rlts) -->
<!-- colnames(stovetype_HS_mean) = colnames(rlts) -->
<!-- colnames(stovetype_LS_mean) = colnames(rlts) -->
<!-- colnames(stovetype_Mix_mean) = colnames(rlts) -->
<!-- colnames(stovetype_PS_mean) = colnames(rlts) -->

<!-- rownames(stovetype_FP_mean) = EAAs_vars -->
<!-- rownames(stovetype_HS_mean) = EAAs_vars -->
<!-- rownames(stovetype_LS_mean) = EAAs_vars -->
<!-- rownames(stovetype_Mix_mean) = EAAs_vars -->
<!-- rownames(stovetype_PS_mean) = EAAs_vars -->

<!-- stovetype_FP_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL"))) -->
<!-- stovetype_HS_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL"))) -->
<!-- stovetype_LS_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL"))) -->
<!-- stovetype_Mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL"))) -->
<!-- stovetype_PS_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL"))) -->

<!-- p_FP <- make_forest_plot(stovetype_FP_mean, "FP (reference)") -->
<!-- p_HS <- make_forest_plot(stovetype_HS_mean, "HS") -->
<!-- p_LS <- make_forest_plot(stovetype_LS_mean, "LS") -->
<!-- p_Mix <- make_forest_plot(stovetype_Mix_mean, "Mix") -->
<!-- p_PS <- make_forest_plot(stovetype_PS_mean, "PS") -->

<!-- plot_list <- list(p_FP, p_HS, p_LS, p_Mix, p_PS) -->

<!-- # grid.arrange(grobs = plot_list,  ncol = 2, -->
<!-- #              top = textGrob(paste("Average EAA of each Fuel Type"), gp=gpar(fontsize=15, fontface = 'bold'))) -->

<!-- for(j in 1:4){ -->
<!--   plot_list[[j]] <- plot_list[[j]] + ylim(c(-7, 8)) -->
<!-- } -->
<!-- grid.arrange(grobs = plot_list,  ncol = 2, -->
<!--              top = textGrob(paste("Average EAA difference of each Stove Type compared to FP"), gp=gpar(fontsize=15, fontface = 'bold'))) -->

<!-- ``` -->

# Linear regression with log transformation

In the following section, we performed regression with equation $$Y = \beta_0 + \beta_1 *log(x)$$
where $Y$ is one of the epigenetic age accelerations, and $X$ is one of the ambient exposure measurements or the urinary measurements.

The estimations of $\beta_1$ with given $Y$ and $log(X)$ are shown below, which can be interpreted as "the mean of Y changes given a one-unit increase in log(X) while holding other variables constant".

## Ambient Exposure
```{r linear_reg_coef_log_ambient, fig.width=10, fig.height=16}
lm_coef_ambient <- data.frame()
lm_coef_ambient_plots <- list()

for(j in 1:length(ambient_exp_vars)){
  plot_df <- data.frame(matrix(NA, ncol = 4))
  for(i in 1:length(EAAs_vars)){
    rlts = preform_regression(Y = EAAs_vars[i], X = ambient_exp_vars[j], log_trans_X = T)
    lm_coef_ambient[i,j] <- rlts[1]
    plot_df[i,] <- rlts[c(2,3,4,5)]
  }
  colnames(plot_df) <- c("coefficient", "cl_lower", "cl_upper", "p_val")
  rownames(plot_df) <- EAAs_vars
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$cl_lower = as.numeric(plot_df$cl_lower)
  plot_df$cl_upper = as.numeric(plot_df$cl_upper)

  pj <- make_forest_plot(plot_df, ambient_exp_label[j])
  lm_coef_ambient_plots[[j]] = pj
}


lm_coef_ambient[9,1] <- c("[P<0.001: ***; P<0.01: **; P<0.05: *]")
colnames(lm_coef_ambient) <- ambient_exp_names
rownames(lm_coef_ambient) <- c(EAAs_names, "")



grid.arrange(grobs = lm_coef_ambient_plots,  ncol = 3,
             top = textGrob(paste("Ambient Exposures (log transformed)"), gp=gpar(fontsize=15, fontface = 'bold')))

for(j in 1:length(ambient_exp_vars)){
  lm_coef_ambient_plots[[j]] <- lm_coef_ambient_plots[[j]] + ylim(c(-2, 2.5))
}
grid.arrange(grobs = lm_coef_ambient_plots,  ncol = 3,
             top = textGrob(paste("Ambient Exposures (log transformed) (same range)"), gp=gpar(fontsize=15, fontface = 'bold')))

cat("X: Ambient Exposure Measurements:")
print(lm_coef_ambient)

```

## Urinary Measurements
```{r linear_reg_coef_log_urine, fig.width=10, fig.height=9}
lm_coef_urine <- data.frame()
lm_coef_urine_plots <- list()

for(j in 1:length(urinary_exp_vars)){
  plot_df <- data.frame(matrix(NA, ncol = 4))
  for(i in 1:length(EAAs_vars)){
    rlts = preform_regression(Y = EAAs_vars[i], X = urinary_exp_vars[j], log_trans_X = T)
    lm_coef_urine[i,j] <- rlts[1]
    plot_df[i,] <- rlts[c(2,3,4,5)]
  }
  colnames(plot_df) <- c("coefficient", "cl_lower", "cl_upper", "p_val")
  rownames(plot_df) <- EAAs_vars
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$cl_lower = as.numeric(plot_df$cl_lower)
  plot_df$cl_upper = as.numeric(plot_df$cl_upper)

  pj <- make_forest_plot(plot_df, urinary_exp_label[j])
  lm_coef_urine_plots[[j]] = pj
}


lm_coef_urine[9,1] <- c("[P<0.001: ***; P<0.01: **; P<0.05: *]")
colnames(lm_coef_urine) <- urinary_exp_names
rownames(lm_coef_urine) <- c(EAAs_names, "")



grid.arrange(grobs = lm_coef_urine_plots,  ncol = 3,
             top = textGrob(paste("Urinary Measurements (log transformed)"), gp=gpar(fontsize=15, fontface = 'bold')))

for(j in 1:length(urinary_exp_vars)){
  lm_coef_urine_plots[[j]] <- lm_coef_urine_plots[[j]] + ylim(c(-2, 2))
}
grid.arrange(grobs = lm_coef_urine_plots,  ncol = 3,
             top = textGrob(paste("Urinary Measurements (log transformed) (same range)"), gp=gpar(fontsize=15, fontface = 'bold')))

cat("X: Urine Measurements:")
print(lm_coef_urine)

```


```{r}
# write.csv(lm_coef_ambient, "../data/2.analysis/lm_coefficients_ambient.csv")
# write.csv(lm_coef_urine, "../data/2.analysis/lm_coefficients_urine.csv")
```

<!-- ```{r} -->
<!-- library(quantreg) -->

<!-- run_quantile_regression <- function(y_var, x_var){ -->
<!--   Y <- y_var -->
<!--   X <- "BMI + Age" -->
<!--   form <- as.formula(paste0(Y,"~", x_var, "+" ,X)) -->

<!--   rqfit <- rq(form, data = df) -->
<!--   return(as.numeric(summary(rqfit,se = "nid")$coefficients[2,])) -->
<!-- } -->

<!-- coef_df <- data.frame() -->
<!-- for(y_var in clocks_vars){ -->
<!--   for (x_var in c(urinary_exp_vars, ambient_exp_vars)) { -->
<!--     coef_df <- rbind(coef_df, c(y_var, x_var, run_quantile_regression(y_var, x_var))) -->
<!--   } -->
<!-- } -->
<!-- names(coef_df) = c("Clock", "Exposure_Measurement", "Ceofficient", "Std.Error", "t_value", "p_value") -->

<!-- cat("Tests with coefficients with p-values < 0.1: \n") -->
<!-- coef_df[coef_df$p_value < 0.1,] -->
<!-- ``` -->


<!-- ## Quantile Regression -->
<!-- ```{r} -->
<!-- library(quantreg) -->

<!-- run_quantile_regression <- function(y_var, x_var){ -->
<!--   Y <- y_var -->
<!--   X <- "BMI + Age" -->
<!--   form <- as.formula(paste0(Y,"~", x_var, "+" ,X)) -->

<!--   rqfit <- rq(form, data = df) -->
<!--   return(as.numeric(summary(rqfit,se = "nid")$coefficients[2,])) -->
<!-- } -->

<!-- coef_df <- data.frame() -->
<!-- for(y_var in clocks_vars){ -->
<!--   for (x_var in c(urinary_exp_vars, ambient_exp_vars)) { -->
<!--     coef_df <- rbind(coef_df, c(y_var, x_var, run_quantile_regression(y_var, x_var))) -->
<!--   } -->
<!-- } -->
<!-- names(coef_df) = c("Clock", "Exposure_Measurement", "Ceofficient", "Std.Error", "t_value", "p_value") -->

<!-- cat("Tests with coefficients with p-values < 0.1: \n") -->
<!-- coef_df[coef_df$p_value < 0.1,] -->
<!-- ``` -->

<!-- ## Quantile Regression -->
<!-- ```{r} -->
<!-- library(quantreg) -->

<!-- run_quantile_regression <- function(y_var, x_var){ -->
<!--   Y <- y_var -->
<!--   X <- "BMI + Age" -->
<!--   form <- as.formula(paste0(Y,"~", x_var, "+" ,X)) -->

<!--   rqfit <- rq(form, data = df) -->
<!--   return(as.numeric(summary(rqfit,se = "nid")$coefficients[2,])) -->
<!-- } -->

<!-- coef_df <- data.frame() -->
<!-- for(y_var in clocks_vars){ -->
<!--   for (x_var in c(urinary_exp_vars, ambient_exp_vars)) { -->
<!--     coef_df <- rbind(coef_df, c(y_var, x_var, run_quantile_regression(y_var, x_var))) -->
<!--   } -->
<!-- } -->
<!-- names(coef_df) = c("Clock", "Exposure_Measurement", "Ceofficient", "Std.Error", "t_value", "p_value") -->

<!-- cat("Tests with coefficients with p-values < 0.1: \n") -->
<!-- coef_df[coef_df$p_value < 0.1,] -->
<!-- ``` -->