---
title: "PASH Analysis"
author: "Seraphina Shi"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
setwd("/Users/seraphinashi/Desktop/PAHS/Xuanwei Study coal and DNA metylation study/Seraphina's folder/PAHS-scripts")

plotFolder <- "../images/03_analysis_v5/"
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=plotFolder,
  cache.path=".cache/",
  duplicate.label="allow"
)

```

```{r}
source('shared_commands.R')
library(table1)
```

```{r}
df <- read.csv("../data/data_cleaned_final.csv") %>% select(-X)
```

```{r}
dup_sbjctD <- df$SbjctD[duplicated(df$SbjctD)]
df_sub1 = df[(df$SbjctD %in% dup_sbjctD & df$Visit == 1) | (! df$SbjctD %in% dup_sbjctD), ]
df_subs <- list("df_sub1" = df_sub1,
                "df_sub2" = df[(df$SbjctD %in% dup_sbjctD & df$Visit == 2) | (! df$SbjctD %in% dup_sbjctD),])
```


```{r define_functions_1}
sig_color <- c("<= 0.001" = "#7CAE00", "<= 0.01" = "#00BFC4", "<= 0.05" = "#C77CFF", "> 0.05" = "black")

make_forest_plot <- function(df_plot, exposure){
  df_plot$sig_level <- as.factor(df_plot$sig_level)
  p <- ggplot(data=df_plot, aes(x=EAAs, y=coefficient, ymin=ci_lower, ymax=ci_upper)) +
    geom_point(shape=24, size=2, aes(col = sig_level, fill = sig_level)) +
    geom_errorbar(shape=24, width=0.3, aes(col = sig_level)) +
    geom_hline(yintercept=0, lty=2,color="red") +  # add a dotted line at x=1 after flip
    # ylim(c(-2,2)) +  # not working
    coord_flip() +  # flip coordinates (puts labels on y axis)
    labs(x=expression(log[2]-PFAS))+
    labs(y="change in 1 year") +
    ggtitle(exposure)+
    theme(plot.title = element_text(size=8.6, hjust = 0.5),
          axis.title.y = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.y = element_text(size=8),
          axis.title.x = element_text(size=8),
          axis.text = element_text(size=7)) +
    scale_color_manual(values = sig_color) +
    scale_fill_manual(values = sig_color) + 
    guides(fill=guide_legend(title="P-value"))

  return(p)
}

make_forest_plot_lg <- function(df, exposure){
  df$sig_level <- as.factor(df$sig_level)
  p <- ggplot(data=df, aes(x=EAAs, y=coefficient, ymin=ci_lower, ymax=ci_upper)) +
    geom_point(shape=24, size=2, aes(col = sig_level, fill = sig_level)) +
    geom_errorbar(shape=24, width=0.3, aes(col = sig_level)) +
    geom_hline(yintercept=0, lty=2,color="red") +  # add a dotted line at x=1 after flip
    # ylim(c(-2,2)) +  # not working
    coord_flip() +  # flip coordinates (puts labels on y axis)
    labs(x=expression(log[2]-PFAS))+
    labs(y="change in 1 year") +
    ggtitle(exposure)+
    theme(plot.title = element_text(size=15, hjust = 0.5),
          axis.title.y = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.y = element_text(size=10),
          axis.title.x = element_text(size=10),
          axis.text = element_text(size=10)) +
    scale_color_manual(values = sig_color) +
    scale_fill_manual(values = sig_color) + 
    guides(fill= guide_legend(title="P-value"),
           col = guide_legend(title="P-value") )

  return(p)
}


options(scipen = 999)
```

```{r define_functions_2}
library(geepack)

preform_gee <- function(Y, X, log_trans_X = F, adjust_cfd = T, dat = df) {
  tmp <- dat[!is.na(dat[,X]),]
  if(log_trans_X) {
    tmp[,X] <- log(tmp[,X])
  }
  
  if(adjust_cfd){
    tmp <- tmp %>% select(var_name_list$ID,
                          var_name_list$EEAs,
                          var_name_list$confounders,
                          unname(X)) %>% na.omit()
  }
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  X_form <- ifelse(adjust_cfd, 
                 paste0(X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(X))

  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") # This says all pairs of responses within a subject are equally correlated
  coef <- summary(geefit)$coefficients[2,1]
  coef_text <- as.character(round(coef,4))
  std <- summary(geefit)$coefficients[2,2]
  std_text <- as.character(round(std,4))
  p_val <- summary(geefit)$coefficients[2,4]
  sig_star <- ifelse(p_val<0.001, "***",
                ifelse(p_val<0.01, "** ",
                       ifelse(p_val<0.05, "*  ", "   ")))
  sig_level <- ifelse(p_val <= 0.001, "<= 0.001",
                ifelse(p_val <= 0.01, "<= 0.01",
                       ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))
  # rlt <- paste0(coef_text, paste0(rep(" ", 7-nchar(coef_text)), collapse = ""),
  #               sig_star,
  #               "(", std_text, paste0(rep(" ", 6-nchar(std_text)), collapse = ""), ")")
  ci_l <- coef - 1.96 * std
  ci_u <- coef + 1.96 * std
  return(c(coef, std, ci_l, ci_u, p_val, sig_level))
}


perform_group_gee <- function(X, adjust_cfd = T, var_name = NA){
  plot_df_list <- list()
  lm_coef_ambient_plot_list <- list()
  for (j in 1:length(X)) {
    exp <- X[j]
    plot_df <- data.frame(matrix(NA, ncol = 6))
    for(i in 1:length(var_name_list$EEAs)){
      rlts = preform_gee(Y = var_name_list$EEAs[i], 
                                X = exp, 
                                adjust_cfd = adjust_cfd, 
                                dat=df)
      plot_df[i,] <- rlts
    }
    colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
    rownames(plot_df) <- var_name_list$EEAs
    plot_df$EAAs <- EAAs_names
    plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
  
    plot_df$coefficient = as.numeric(plot_df$coefficient)
    plot_df$ci_lower = as.numeric(plot_df$ci_lower)
    plot_df$ci_upper = as.numeric(plot_df$ci_upper)
  
    if(sum(is.na(var_name))>0){
      lm_coef_ambient_plot <- make_forest_plot(plot_df, exp)
    } else {
      lm_coef_ambient_plot <- make_forest_plot(plot_df, var_name[j])
    }
    
    plot_df_list[[exp]] <- plot_df
    lm_coef_ambient_plot_list[[exp]] = lm_coef_ambient_plot
  }
  return(list(plot_df_list, lm_coef_ambient_plot_list))
}



preform_gee_exposure_prototype <- function(X, Y, adjust_cfd = T, log_trans_X=F) {
  tmp <- df
  for(i in 1:length(X)){
    X_i <- X[i]
    tmp <- tmp[!is.na(tmp[,X_i]),]
    if(log_trans_X) {
      tmp[,X_i] <- log(tmp[,X_i])
    }
  }
  
  X_form <- paste(X, collapse  = " + ")

  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X_form, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~", X_form))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") # This says all pairs of responses within a subject are equally correlated


  coefs <- as.data.frame(summary(geefit)$coefficients)
  colnames(coefs) <- c("coefficient", "std", "t_val", "pval")

  coefs <- coefs %>%
    select(coefficient, std, pval) %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(pval <= 0.001, "<= 0.001",
                        ifelse(pval <= 0.01, "<= 0.01",
                               ifelse(pval <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, pval, sig_level)


  colnames(coefs) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  
  return(coefs)
}

```
In this file, we presented the results from analyses of exposures against computed epigenetic age accelerations (EAA) calculated using DNA methylation data by different methods. For each exposure, we conducted both primary analyses (likelihood ratio tests and generalized estimating equations (GEE), adjusted for confounders) and sensitivity analyses (likelihood ratio tests and/or generalized estimating equations (GEE) limited to certain fuel users, not adjusted for confounders). In addition to what's included in the analysis plan, we also analyzed ambient and urinary exposures. 



# 1. Table with summarize info [table 1]
## 1.1. Description of study population (first visit for all objects)
There are `r nrow(df)` visits with corresponding epigenetic ages available among `r length(unique(df$SbjctD))` female subjects. For these `r length(unique(df$SbjctD))` subjects, `r length(unique(df$SbjctD)) - length(dup_sbjctD)` have been visited once and `r length(dup_sbjctD)` have been visited twice. 

The following tables summarize all the information of the first visit of these `r length(unique(df$SbjctD))` subjects.

### Baseline characteristics (confounders)
```{r}
library(table1)
table1(~ Age + BMI + factor(county) + factor(ses)  + factor(edu), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```


### Epigenetic ages
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$clocks, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

### Epigenetic ages accelarations
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$EEAs, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

### Fuel/stove type exposures
```{r}
table1(as.formula(paste(" ~ ", paste(paste0("factor(", c(var_name_list$fuel_exp, var_name_list$stove_exp),")"), 
                                     collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

### 5MC exposures
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```


```{r}
print("Pearson pair-wise correlation:")
cor(df_sub1 %>%  select(var_name_list$PCH_5MC) %>% na.omit(), 
    method = "pearson")
```
```{r}
print("Spearman pair-wise correlation:")
cor(df_sub1 %>%  select(var_name_list$PCH_5MC) %>% na.omit(), 
    method = "spearman")
```

### Cluster-based exposures
#### clusCUR6
Clusters based on model-based exposure estimates at or shortly before the visit
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUR6, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

#### clusCHLD5
Clusters based on model-based exposure estimates accrued before age 18 
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCHLD5, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

#### clusCUM6
Clusters based on model-based lifetime exposure estimates 
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUM6, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

#### clusMEAS6
Clusters based on pollutant measurements
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusMEAS6, collapse = " + "))), 
       data = df_sub1)
```

#### clusURI5
Clusters based on urinary biomarkers
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusURI5, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```



### Ambient exposures
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$ambient_exp, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

### Urinary biomarkers
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$urinary_exp, collapse = " + "))), 
       data = df_sub1,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```




# 2.1. Current fuel type 
## Primary analysis
### GEE (with confounders, Smokeless ref)
The numbers of observations with each current fuel type:
```{r}
table(df$curFuel)
```

In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky}) + \beta_2 * I(\text{Wood_and_or_Plant})  \\
  & + \beta_3 * county + \beta_4 * BMI + \beta_5 * ses + \beta_6 * edu + \epsilon
\end{aligned}
$$


Results:
```{r}
preform_gee_curFuel <- function(Y, adjust_cfd = T) {
  X = "curFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp$curFuel = factor(tmp$curFuel, levels = c("Smokeles","Smoky","Wood_and_or_Plant"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Smokeless (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "curFuelSmoky"] = "Smoky"
  rownames(coefs)[rownames(coefs) == "curFuelWood_and_or_Plant"] = "Wood_and_or_Plant"  

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_curFuel_mix_org, fig.width=7, fig.height=3}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smokeless_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_woodplant_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_curFuel(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_smokeless_mean[i,] <- rlts[1,]
    fueltype_smoky_mean[i,] <- rlts[2,]
    fueltype_woodplant_mean[i,] <- rlts[3,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_smokeless_mean) = colnames(rlts)
colnames(fueltype_woodplant_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = var_name_list$EEAs
rownames(fueltype_smokeless_mean) = var_name_list$EEAs
rownames(fueltype_woodplant_mean) = var_name_list$EEAs

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_woodplant_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

coefs_cur <- fueltype_smoky_mean

p1_smokeles <- make_forest_plot_lg(fueltype_smokeless_mean, "Smokeless (reference)")
p1_smoky <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")
p1_woodplant <- make_forest_plot_lg(fueltype_woodplant_mean, "Wood_and_or_Plant")

plot_list <- list(p1_smoky, p1_woodplant)

for(j in 1:2){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-8, 10))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Average EAA difference of each current Fuel Type\n compared to Smokeless Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to Smokeless fuel yupes:")
print(fueltype_smoky_mean %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```

```{r}
cat("The estimated average EAA differences of Wood and/or Plant fuel type to Smokeless fuel yupes:")
print(fueltype_woodplant_mean %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```


### GEE (with confounders, Other ref)
The numbers of observations with each current fuel type:
```{r}
df <- df %>% mutate(curFuel_1 = ifelse(curFuel=="Smoky", "Smoky", "Other"))
table(df$curFuel_1)
```

In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \beta_3 * county + \beta_4 * BMI + \beta_5 * ses + \beta_6 * edu + \epsilon
\end{aligned}
$$


Results:
```{r}
preform_gee_curFuel_1 <- function(Y, adjust_cfd = T) {
  X = "curFuel_1"
  tmp <- df[!is.na(df[,X]),]

  tmp$curFuel_1 = factor(tmp$curFuel_1, levels = c("Other","Smoky"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[1] = "Other (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "curFuelSmoky"] = "Smoky"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_curFuel_mix_org_other, fig.width=4, fig.height=3}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_other_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_curFuel_1(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_other_mean[i,] <- rlts[1,]
    fueltype_smoky_mean[i,] <- rlts[2,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_other_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = var_name_list$EEAs
rownames(fueltype_other_mean) = var_name_list$EEAs

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_other_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

fueltype_smoky_mean_4 <- fueltype_smoky_mean

p1_other <- make_forest_plot_lg(fueltype_other_mean, "Other (reference)")
p1_smoky <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky \n (compared to Other fuel types)")

p1_smoky

```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to Other fuel types:")
print(fueltype_smoky_mean_4 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```




### GEE (with confounders, Wood_and_or_Plant ref)
The numbers of observations with each current fuel type:
```{r}
table(df$curFuel)
```

In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky}) + \beta_2 * I(\text{Wood_and_or_Plant})  \\
  & + \beta_3 * county + \beta_4 * BMI + \beta_5 * ses + \beta_6 * edu + \epsilon
\end{aligned}
$$


Results:
```{r}
preform_gee_curFuel_3 <- function(Y, adjust_cfd = T) {
  X = "curFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp$curFuel = factor(tmp$curFuel, levels = c("Wood_and_or_Plant", "Smoky", "Smokeles"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[1] = "Wood_and_or_Plant (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "curFuelSmoky"] = "Smoky"
  rownames(coefs)[rownames(coefs) == "curFuelSmokeles"] = "Smokeles"  

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_curFuel_mix_org_1, fig.width=7, fig.height=3}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smokeless_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_woodplant_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_curFuel_3(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_smokeless_mean[i,] <- rlts[3,]
    fueltype_smoky_mean[i,] <- rlts[2,]
    fueltype_woodplant_mean[i,] <- rlts[1,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_smokeless_mean) = colnames(rlts)
colnames(fueltype_woodplant_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = var_name_list$EEAs
rownames(fueltype_smokeless_mean) = var_name_list$EEAs
rownames(fueltype_woodplant_mean) = var_name_list$EEAs

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_woodplant_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))


p1_woodplant <- make_forest_plot_lg(fueltype_smokeless_mean, "Wood_and_or_Plant (reference)")
p1_smoky <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")
p1_smokeles <- make_forest_plot_lg(fueltype_woodplant_mean, "Smokeless")

plot_list <- list(p1_smoky, p1_smokeles)

for(j in 1:2){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-8, 10))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Average EAA difference of each current Fuel Type\n compared to Wood_and_or_Plant Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```



### GEE (with confounders, Smokeless_PureWood ref)
The numbers of observations with each current fuel type:
```{r}
table(df$curFuel_detail)
```

```{r}
table(df$curFuel_detail, df$childFuel)
```

```{r}
table(df$curFuel_detail, df$cumFuel)
```

```{r}
pure_wood_id <- df$SbjctD[df$curFuel_detail=="Wood" & df$childFuel=="Wood"]
pure_wood_id <- pure_wood_id[!is.na(pure_wood_id)]

df$curFuel_detail[df$SbjctD %in% pure_wood_id] <- "Smokeless_PureWood"
df$curFuel_detail[df$curFuel_detail %in% "Smokeles"] <- "Smokeless_PureWood"

table(df$curFuel_detail)
```


Results:
```{r}
preform_gee_curFuel_5 <- function(Y, adjust_cfd = T) {
  X = "curFuel_detail"
  tmp <- df[!is.na(df[,X]),]
  
  tmp$curFuel_detail = factor(tmp$curFuel_detail, levels = c("Smokeless_PureWood","Smoky","Wood", "Plant"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3,4), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[1] = "Smokeless_PureWood (reference/intercept)"
  rownames(coefs)[2] = "Smoky"
  rownames(coefs)[3] = "Wood"  
  rownames(coefs)[4] = "Plant"  

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_curFuel_mix_org_5, fig.width=7, fig.height=5}
fueltype_ref_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_wood_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_plan_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_curFuel_5(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_ref_mean[i,] <- rlts[1,]
    fueltype_smoky_mean[i,] <- rlts[2,]
    fueltype_wood_mean[i,] <- rlts[3,]
    fueltype_plan_mean[i,] <- rlts[4,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_ref_mean) = colnames(rlts)
colnames(fueltype_wood_mean) = colnames(rlts)
colnames(fueltype_plan_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = var_name_list$EEAs
rownames(fueltype_ref_mean) = var_name_list$EEAs
rownames(fueltype_wood_mean) = var_name_list$EEAs
rownames(fueltype_plan_mean) = var_name_list$EEAs

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_ref_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_wood_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_plan_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")
p2 <- make_forest_plot_lg(fueltype_wood_mean, "Wood")
p3 <- make_forest_plot_lg(fueltype_plan_mean, "Plan")

plot_list <- list(p1, p2, p3)

for(j in 1:3){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-10, 10))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Average EAA difference of each current Fuel Type\n compared to Smokeless_PureWood Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```






# 2.2. Birth fuel type 
## Primary analysis
### GEE (with confounders, Smokeless ref)
The numbers of observations with each birth fuel type:
```{r}
table(df$brthFuel)
```

Results:
```{r}
preform_gee_brthFuel <- function(Y, adjust_cfd = T) {
  X = "brthFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$brthFuel != "Outside of XW/FY",]
  
  tmp$brthFuel = factor(tmp$brthFuel, levels = c("Smokeles","Smoky","Wood", "Mix"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[1:5, c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_brthFuel_mix, fig.width=7, fig.height=5}
coefs1 <- data.frame(matrix(NA, ncol = 6))
coefs2 <- data.frame(matrix(NA, ncol = 6))
coefs3 <- data.frame(matrix(NA, ncol = 6))
coefs4 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_brthFuel(Y = var_name_list$EEAs[i], adjust_cfd = T)
    coefs1[i,] <- rlts[1,]
    coefs2[i,] <- rlts[2,]
    coefs3[i,] <- rlts[3,]
    coefs4[i,] <- rlts[4,]
}

colnames(coefs1) = colnames(rlts)
colnames(coefs2) = colnames(rlts)
colnames(coefs3) = colnames(rlts)
colnames(coefs4) = colnames(rlts)

rownames(coefs1) = var_name_list$EEAs
rownames(coefs2) = var_name_list$EEAs
rownames(coefs3) = var_name_list$EEAs
rownames(coefs4) = var_name_list$EEAs


coefs1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coefs2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coefs3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coefs4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coefs1, "Smokeless (reference)")
p2 <- make_forest_plot_lg(coefs2, "Smoky")
p3 <- make_forest_plot_lg(coefs3, "Wood")
p4 <- make_forest_plot_lg(coefs4, "Mix")

plot_list <- list(p2, p3, p4)

for(j in 1:3){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-10, 13))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Average EAA difference of each birth Fuel Type\n compared to Smokeless Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to Smokeless fuel yupes:")
print(coefs2 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```

```{r}
cat("The estimated average EAA differences of Wood fuel type to Smokeless fuel yupes:")
print(coefs3 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```

```{r}
cat("The estimated average EAA differences of Mix fuel type to Smokeless fuel yupes:")
print(coefs4 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```

### GEE (with confounders, Mix ref)
The numbers of observations with each birth fuel type:
```{r}
table(df$brthFuel)
```

Results:
```{r}
preform_gee_brthFuel_2 <- function(Y, adjust_cfd = T) {
  X = "brthFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[tmp$brthFuel != "Outside of XW/FY",]
  
  tmp$brthFuel = factor(tmp$brthFuel, levels = c("Mix","Smoky","Smokeles","Wood"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[1:5, c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_brthFuel_mix_1, fig.width=7, fig.height=5}
coefs1 <- data.frame(matrix(NA, ncol = 6))
coefs2 <- data.frame(matrix(NA, ncol = 6))
coefs3 <- data.frame(matrix(NA, ncol = 6))
coefs4 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_brthFuel(Y = var_name_list$EEAs[i], adjust_cfd = T)
    coefs1[i,] <- rlts[1,]
    coefs2[i,] <- rlts[2,]
    coefs3[i,] <- rlts[3,]
    coefs4[i,] <- rlts[4,]
}

colnames(coefs1) = colnames(rlts)
colnames(coefs2) = colnames(rlts)
colnames(coefs3) = colnames(rlts)
colnames(coefs4) = colnames(rlts)

rownames(coefs1) = var_name_list$EEAs
rownames(coefs2) = var_name_list$EEAs
rownames(coefs3) = var_name_list$EEAs
rownames(coefs4) = var_name_list$EEAs


coefs1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coefs2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coefs3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coefs4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

coefs_brth <- coefs2

p1 <- make_forest_plot_lg(coefs1, "Mix (reference)")
p2 <- make_forest_plot_lg(coefs2, "Smoky")
p3 <- make_forest_plot_lg(coefs3, "Smokeless")
p4 <- make_forest_plot_lg(coefs4, "Wood")

plot_list <- list(p2, p3, p4)

for(j in 1:3){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-10, 13))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Average EAA difference of each birth Fuel Type\n compared to Mix Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to Smokeless fuel yupes:")
print(coefs2 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```

```{r}
cat("The estimated average EAA differences of Wood fuel type to Smokeless fuel yupes:")
print(coefs3 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```

```{r}
cat("The estimated average EAA differences of Mix fuel type to Smokeless fuel yupes:")
print(coefs4 %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```


### GEE (with confounders, Other ref)
The numbers of observations with each current fuel type:
```{r}
df <- df %>% mutate(brthFuel_1 = ifelse(brthFuel=="Smoky", "Smoky", "Other"))
table(df$brthFuel_1[df$brthFuel != "Outside of XW/FY"])
```

In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between current fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \beta_3 * county + \beta_4 * BMI + \beta_5 * ses + \beta_6 * edu + \epsilon
\end{aligned}
$$


Results:
```{r}
preform_gee_brthFuel_1 <- function(Y, adjust_cfd = T) {
  X = "brthFuel_1"
  tmp <- df[!is.na(df[,X]),]
  tmp <- tmp[tmp$brthFuel != "Outside of XW/FY", ]
  tmp$brthFuel_1 = factor(tmp$brthFuel_1, levels = c("Other","Smoky"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[1] = "Other (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "curFuelSmoky"] = "Smoky"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_brthFuel_mix_org_other, fig.width=4, fig.height=3}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_other_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_brthFuel_1(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_other_mean[i,] <- rlts[1,]
    fueltype_smoky_mean[i,] <- rlts[2,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_other_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = var_name_list$EEAs
rownames(fueltype_other_mean) = var_name_list$EEAs

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_other_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1_other <- make_forest_plot_lg(fueltype_other_mean, "Other (reference)")
p1_smoky <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky \n (compared to Other fuel types)")

p1_smoky

```

```{r}
cat("The estimated average EAA differences of Smoky fuel type to Other fuel types:")
print(fueltype_smoky_mean %>% 
        mutate(coefficient = round(coefficient, 4),
               std = round(std, 4),
               ci_lower = round(ci_lower, 4),
               ci_upper = round(ci_upper, 4),
               p_val = round(p_val, 6)) %>%
        select(-EAAs))
```







# 2.3. Childhood  fuel type 

## Primary analysis
## GEE (with confounders, Smokeless ref)
The numbers of observations with each current fuel type:
```{r}
table(df$childFuel)
```
In this section, we perform the generalized estimating equations (GEE) to evaluate the association between current fuel type and the Grim Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky}) + \beta_2 * I(\text{Wood}) + \beta_3 * I(\text{Mix})  \\
  & + \beta_4 * county + \beta_5 * BMI + \beta_6 * ses + \beta_7 * edu + \epsilon
\end{aligned}
$$

Results:
```{r}
preform_gee_childFuel <- function(Y, adjust_cfd = T) {
  X = "childFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp$childFuel = factor(tmp$childFuel, levels = c("Smokeles", "Smoky","Wood", "Mix"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3,4), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Smokelees (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "childFuelSmoky"] = "Smoky"
  rownames(coefs)[rownames(coefs) == "childFuelWood"] = "Wood"
  rownames(coefs)[rownames(coefs) == "childFuelMix"] = "Mix"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_childFuel_mix_org, fig.width=7, fig.height=5}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_wood_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smokeless_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_mix_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_childFuel(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_smoky_mean[i,] <- rlts[2,]
    fueltype_wood_mean[i,] <- rlts[3,]
    fueltype_smokeless_mean[i,] <- rlts[1,]
    fueltype_mix_mean[i,] <- rlts[4,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_wood_mean) = colnames(rlts)
colnames(fueltype_smokeless_mean) = colnames(rlts)
colnames(fueltype_mix_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_wood_mean) = EAAs_vars
rownames(fueltype_smokeless_mean) = EAAs_vars
rownames(fueltype_mix_mean) = EAAs_vars

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_wood_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

fueltype_smoky_mean_2 <- fueltype_smoky_mean

p2 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")
p3 <- make_forest_plot_lg(fueltype_wood_mean, "Wood")
p1 <- make_forest_plot_lg(fueltype_smokeless_mean, "Smokeless (reference)")
p4 <- make_forest_plot_lg(fueltype_mix_mean, "Mix")

plot_list <- list(p2, p3, p4)

for(j in 1:3){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-7.5, 10))
}

grid.arrange(grobs = plot_list,  ncol = 2, 
             layout_matrix = rbind(c(1,2),
                                   c(3,NA)),
             top = textGrob(paste("Average EAA difference of each Childhood Fuel Type\n compared to Smokeless Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
cat("The estimated average EAA difference of childhood fuel type, smoky compared to smokeless coal:")
print(fueltype_smoky_mean_2 %>% mutate(coefficient = round(coefficient, 4),                        
                       std = round(std, 4),                        
                       ci_lower = round(ci_lower, 4),                        
                       ci_upper = round(ci_upper, 4),                        
                       p_val = round(p_val, 6)))
```

```{r}
cat("The estimated average EAA difference of childhood fuel type, wood compared to smokeless coal:")
print(fueltype_wood_mean %>% mutate(coefficient = round(coefficient, 4),                        
                       std = round(std, 4),                        
                       ci_lower = round(ci_lower, 4),                        
                       ci_upper = round(ci_upper, 4),                        
                       p_val = round(p_val, 6)))
```

```{r}
cat("The estimated average EAA difference of childhood fuel type, mix compared to smokeless coal:")
print(fueltype_mix_mean %>% mutate(coefficient = round(coefficient, 4),                        
                       std = round(std, 4),                        
                       ci_lower = round(ci_lower, 4),                        
                       ci_upper = round(ci_upper, 4),                        
                       p_val = round(p_val, 6)))
```


## GEE (with confounders, Other ref)
The numbers of observations with each current fuel type:
```{r}
df <- df %>% mutate(childFuel_1 = ifelse(childFuel=="Smoky", "Smoky", "Other"))
table(df$childFuel_1)
```
In this section, we perform the generalized estimating equations (GEE) to evaluate the association between current fuel type and the Grim Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky})  \\
  & + \beta_4 * county + \beta_5 * BMI + \beta_6 * ses + \beta_7 * edu + \epsilon
\end{aligned}
$$

Results:
```{r}
preform_gee_childFuel_1 <- function(Y, adjust_cfd = T) {
  X = "childFuel_1"
  tmp <- df[!is.na(df[,X]),]
  
  tmp$childFuel_1 = factor(tmp$childFuel_1, levels = c("Other", "Smoky"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[1] = "Other (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "childFuelSmoky"] = "Smoky"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_childFuel_mix_org_other_2, fig.width=4, fig.height=3}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_other_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_childFuel_1(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_other_mean[i,] <- rlts[1,]
    fueltype_smoky_mean[i,] <- rlts[2,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_other_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_other_mean) = EAAs_vars

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_other_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p2 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")
p1 <- make_forest_plot_lg(fueltype_smokeless_mean, "Other (reference)")

plot_list <- list(p2, p3, p4)

p2

```

```{r}
cat("The estimated average EAA difference of childhood fuel type, smoky compared to other coal:")
print(fueltype_smoky_mean %>% mutate(coefficient = round(coefficient, 4),                        
                       std = round(std, 4),                        
                       ci_lower = round(ci_lower, 4),                        
                       ci_upper = round(ci_upper, 4),                        
                       p_val = round(p_val, 6)))
```



## GEE (with confounders, Smokeless_Wood ref)
The numbers of observations with each current fuel type:
```{r}
df <- df %>% mutate(childFuel_2 = ifelse(childFuel=="Smoky", "Smoky", 
                                           ifelse(childFuel=="Mix", "Mix", "Smokeless_Wood")))
table(df$childFuel_2)
```

Results:
```{r}
preform_gee_childFuel_2 <- function(Y, adjust_cfd = T) {
  X = "childFuel_2"
  tmp <- df[!is.na(df[,X]),]
  
  tmp$childFuel_2 = factor(tmp$childFuel_2, levels = c("Smokeless_Wood", "Smoky", "Mix"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[1] = "Smokeless_Wood (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "childFuelSmoky"] = "Smoky"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_childFuel_mix_org_other_1, fig.width=7, fig.height=5}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_mix_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_childFuel_2(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_mix_mean[i,] <- rlts[3,]
    fueltype_smoky_mean[i,] <- rlts[2,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_mix_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_mix_mean) = EAAs_vars

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

fueltype_smoky_mean_5 <- fueltype_smoky_mean

p2 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")
p1 <- make_forest_plot_lg(fueltype_smokeless_mean, "Mix")

plot_list <- list(p1, p2)

for(j in 1:2){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-12, 12))
}

grid.arrange(grobs = plot_list,  ncol = 2, 
             top = textGrob(paste("Average EAA difference of each Childhood Fuel Type\n compared to Smokeless_Wood Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
cat("The estimated average EAA difference of childhood fuel type, smoky compared to Smokeless_Wood coal:")
print(fueltype_smoky_mean %>% mutate(coefficient = round(coefficient, 4),                        
                       std = round(std, 4),                        
                       ci_lower = round(ci_lower, 4),                        
                       ci_upper = round(ci_upper, 4),                        
                       p_val = round(p_val, 6)))
```




## GEE (with confounders, Mix ref)
The numbers of observations with each current fuel type:
```{r}
df <- df %>% mutate(childFuel_3 = ifelse(childFuel=="Smoky", "Smoky", 
                                           ifelse(childFuel=="Mix", "Mix", "Smokeless_Wood")))
df$childFuel_3 = factor(df$childFuel_3, levels = c( "Mix", "Smoky", "Smokeless_Wood"))

table(df$childFuel_3)
```


Results:
```{r}
preform_gee_childFuel_3 <- function(Y, adjust_cfd = T) {
  X = "childFuel_3"
  tmp <- df[!is.na(df[,X]),]
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[1] = "Mix (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "childFuelSmoky"] = "Smoky"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_childFuel_mix_org_other, fig.width=7, fig.height=3}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 6))
fueltype_smokeless_wood_mean <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_childFuel_3(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_smokeless_wood_mean[i,] <- rlts[3,]
    fueltype_smoky_mean[i,] <- rlts[2,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_smokeless_wood_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = EAAs_vars
rownames(fueltype_smokeless_wood_mean) = EAAs_vars

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_smokeless_wood_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

coefs_child <- fueltype_smoky_mean

p2 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")
p1 <- make_forest_plot_lg(fueltype_smokeless_wood_mean, "smokeless_wood")

plot_list <- list(p1, p2)

for(j in 1:2){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-5, 5))
}

grid.arrange(grobs = plot_list,  ncol = 2, 
             top = textGrob(paste("Average EAA difference of each Childhood Fuel Type\n compared to Mix Coal"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
cat("The estimated average EAA difference of childhood fuel type, smoky compared to Mix coal:")
print(fueltype_smoky_mean %>% mutate(coefficient = round(coefficient, 4),                        
                       std = round(std, 4),                        
                       ci_lower = round(ci_lower, 4),                        
                       ci_upper = round(ci_upper, 4),                        
                       p_val = round(p_val, 6)))
```



```{r}
cat("The estimated average EAA difference of childhood fuel type, smokeless_wood compared to Mix coal:")
print(fueltype_smokeless_wood_mean %>% mutate(coefficient = round(coefficient, 4),                        
                       std = round(std, 4),                        
                       ci_lower = round(ci_lower, 4),                        
                       ci_upper = round(ci_upper, 4),                        
                       p_val = round(p_val, 6)))
```

# 2.4. Cumulative lifetime fuel type 


## Primary analysis
### GEE (with confounders, Mix ref)

```{r}
table(df$cumFuel)
```
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between cumulative fuel type and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * I(\text{Smoky}) + \beta_2 * I(\text{Wood}) \\
  & + \beta_3 * county + \beta_4 * BMI + \beta_5 * ses + \beta_6 * edu + \epsilon
\end{aligned}
$$


Results:
```{r}
preform_gee_cumFuel <- function(Y, adjust_cfd = T) {
  X = "cumFuel"
  tmp <- df[!is.na(df[,X]),]
  
  tmp <- tmp[! tmp$cumFuel %in% c("Wood", "Smokeles"), ]
  
  tmp$cumFuel = factor(tmp$cumFuel, levels = c("Mix",  "Smoky"))
  
  form <- ifelse(adjust_cfd, 
                 paste0(Y," ~ ", X, " + ", paste(var_name_list$confounders[-1], collapse  = " + ")),
                 paste0(Y,"~",X))
  form <- as.formula(form)
  
  print(paste("Fitting with", nrow(tmp), "observations."))
  geefit <- geeglm(formula = form, 
                data = tmp,
                id = SbjctD,
                family = gaussian,
                corstr = "exchangeable") 

  coefs <- as.data.frame(summary(geefit)$coefficients)[c(1,2,3), c(1,2,4)]
  colnames(coefs) <- c("coefficient", "std", "p_val")
  rownames(coefs)[rownames(coefs) == "(Intercept)"] = "Mix (reference/intercept)"
  rownames(coefs)[rownames(coefs) == "cumFuelSmoky"] = "Smoky"

  coefs <- coefs %>%
    mutate(ci_lower = coefficient - 1.96 * std,
           ci_upper = coefficient + 1.96 * std,
           sig_level = ifelse(p_val <= 0.001, "<= 0.001",
                        ifelse(p_val <= 0.01, "<= 0.01",
                               ifelse(p_val <= 0.05, "<= 0.05", "> 0.05")))) %>% 
    select(coefficient, std, ci_lower, ci_upper, p_val, sig_level)
  return(coefs)
}
```

```{r gee_cumFuel_mix}
fueltype_smoky_mean <- data.frame(matrix(NA, ncol = 7))
fueltype_mix_mean <- data.frame(matrix(NA, ncol = 7))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_cumFuel(Y = var_name_list$EEAs[i], adjust_cfd = T)
    fueltype_mix_mean[i,] <- rlts[1,]
    fueltype_smoky_mean[i,] <- rlts[2,]
}

colnames(fueltype_smoky_mean) = colnames(rlts)
colnames(fueltype_mix_mean) = colnames(rlts)

rownames(fueltype_smoky_mean) = var_name_list$EEAs
rownames(fueltype_mix_mean) = var_name_list$EEAs

fueltype_smoky_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
fueltype_mix_mean$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

coefs_cum <- fueltype_smoky_mean

p2 <- make_forest_plot_lg(fueltype_smoky_mean, "Smoky")

p2

```

```{r}
cat("The estimated average  EAA difference of cumulative fuel type, smoky compared to mix coal:")
print(fueltype_smoky_mean)
```



# 2.0. All plots together
Current: smoky vs smokeless
Birth: smoky vs mix
Childhood: smoky vs mix
Cumulative: smoky vs mix
```{r gee_FuelTypes_4periods, fig.width=12, fig.height=3}
coefs_cur$exposure_time = "Current Exposure"
coefs_brth$exposure_time = "Birth Exposure"
coefs_child$exposure_time = "Childhood Exposure"
coefs_cum$exposure_time = "Cumulative Exposure" 
coefs_cum <- coefs_cum[, !is.na(colnames(coefs_cum))]

coefs <- rbind(coefs_cur, coefs_brth, coefs_child, coefs_cum)

coefs$exposure_time = factor(coefs$exposure_time, levels = c("Current Exposure","Birth Exposure", "Childhood Exposure", "Cumulative Exposure"))
make_forest_plot_lg(coefs, "Average EAA Difference of Smoky Fuel Type Compared to [Smokeless,Mix,Mix,Mix] Fuel Type") +
  ylim(-8,8) + 
  facet_wrap(~ exposure_time, ncol=4)


# smoky vs smokeless
p1 <- make_forest_plot_lg(coefs_cur, "") + 
  ylim(-6,7) +
  labs(y = " ")
ggsave(file=sprintf("%s/fuelType_current.png",plotFolder), p1, width = 4.5, height = 3, dpi = 1200) 


# smoky vs mix
p2 <- make_forest_plot_lg(coefs_brth, "") + 
  ylim(-6,7) 
ggsave(file=sprintf("%s/fuelType_birth.png",plotFolder), p2, width = 4.5, height = 3, dpi = 1200) 


# smoky vs mix
p3 <- make_forest_plot_lg(coefs_child, "") + 
  ylim(-6,7) 
ggsave(file=sprintf("%s/fuelType_childhood.png",plotFolder), p3, width = 4.5, height = 3, dpi = 1200) 


# smoky vs mix
p4 <- make_forest_plot_lg(coefs_cum, "") + 
  ylim(-6,7) +
  labs(y = " ")
ggsave(file=sprintf("%s/fuelType_cumulative.png",plotFolder), p4, width = 4.5, height = 3, dpi = 1200) 
```




# 3.1. Clusters based on model-based exposure estimates at or shortly before the visit (clusCUR6)
The file LEX_clusCUR6.csv has information on current pollutant exposures, obtained for the 2 years preceding the visit. To reduce multi-collinearity between exposures, exposure prototypes were derived based on hierarchical cluster analysis in combination followed by principal components analysis. These estimates are available for 6 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:

CUR6_BC_PAH6  Black carbon (BC) and 6 PAHs   
CUR6_PAH31  a large cluster of 31 PAHs  
CUR6_NkF  NkF only  
CUR6_PM_RET  Particulate matter (PM) and retene  
CUR6_NO2  NO2 only  
CUR6_SO2  SO2 only    



## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUR6, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis

### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUR6_mix, fig.width=8, fig.height=7}
rlts_gee_1 <- perform_group_gee(X = var_name_list$cluster_exp$clusCUR6, adjust_cfd = T)

for(j in 1:length(rlts_gee_1[[2]])){
  rlts_gee_1[[2]][[j]] <- rlts_gee_1[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee_1[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee_1[[1]])){
  rlts_gee_1[[1]][[j]] <- rlts_gee_1[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee_1[[1]])
```


### GEE (with confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUR6_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUR6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs

rlts_gee_1_mutually_adj <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(rlts_gee_1_mutually_adj) <- var_name_list$cluster_exp$clusCUR6

coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CUR6_BC_PAH6")
p2 <- make_forest_plot_lg(coef2, "CUR6_PAH31")
p3 <- make_forest_plot_lg(coef3, "CUR6_NkF")
p4 <- make_forest_plot_lg(coef4, "CUR6_PM_RET")
p5 <- make_forest_plot_lg(coef5, "CUR6_NO2")
p6 <- make_forest_plot_lg(coef6, "CUR6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUR6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```

## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUR6_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusCUR6, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee[[1]])
```




### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  &  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUR6_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUR6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CUR6_BC_PAH6")
p2 <- make_forest_plot_lg(coef2, "CUR6_PAH31")
p3 <- make_forest_plot_lg(coef3, "CUR6_NkF")
p4 <- make_forest_plot_lg(coef4, "CUR6_PM_RET")
p5 <- make_forest_plot_lg(coef5, "CUR6_NO2")
p6 <- make_forest_plot_lg(coef6, "CUR6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on current pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUR6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```


# 3.2. Clusters based on model-based exposure estimates accrued before age 18 (clusCHLD5)
The file LEX_clusCHLD5.csv has information on estimated pollutant exposures during early childhood. Estimates are available for 5 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:

CHLD5_X7  a cluster of 7 air pollutants  
CHLD5_X33  a large cluster of 33 air pollutants  
CHLD5_NkF  NkF only  
CHLD5_NO2  NO2 only  
CHLD5_SO2  SO2 only    


## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCHLD5, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCHLD5_mix, fig.width=8, fig.height=7}
rlts_gee_2 <- perform_group_gee(X = var_name_list$cluster_exp$clusCHLD5, adjust_cfd = T)

for(j in 1:length(rlts_gee_2[[2]])){
  rlts_gee_2[[2]][[j]] <- rlts_gee_2[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee_2[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee_2[[1]])){
  rlts_gee_2[[1]][[j]] <- rlts_gee_2[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee_2[[1]])
```




### GEE (with confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X7 +  \beta_2 X33 + \beta_3 NkF + \beta_4 NO2 + \beta_5 SO2  \\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_9 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCHLD5_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCHLD5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

rlts_gee_2_mutually_adj <- list(coef1, coef2, coef3, coef4, coef5)
names(rlts_gee_2_mutually_adj) <- var_name_list$cluster_exp$clusCHLD5

p1 <- make_forest_plot_lg(coef1, "CHLD5_X7")
p2 <- make_forest_plot_lg(coef2, "CHLD5_X33")
p3 <- make_forest_plot_lg(coef3, "CHLD5_NkF")
p4 <- make_forest_plot_lg(coef4, "CHLD5_NO2")
p5 <- make_forest_plot_lg(coef5, "CHLD5_SO2")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusCHLD5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```



## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCHLD5_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusCHLD5, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-4, 4))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee[[1]])
```





### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X7 +  \beta_2 X33 + \beta_3 NkF + \beta_4 NO2 + \beta_5 SO2  \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCHLD5_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCHLD5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CHLD5_X7")
p2 <- make_forest_plot_lg(coef2, "CHLD5_X33")
p3 <- make_forest_plot_lg(coef3, "CHLD5_NkF")
p4 <- make_forest_plot_lg(coef4, "CHLD5_NO2")
p5 <- make_forest_plot_lg(coef5, "CHLD5_SO2")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-4,4))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on childhood pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusCHLD5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```




# 3.3. Clusters based on model-based lifetime exposure estimates (clusCUM6)
The file LEX_clus CUM6.csv has information on estimated cumulative pollutant exposures during the lifecourse. Estimates are available for 6 different prototypes (cluster variables) for a total of 161 subjects and 211 visits. The prototypes are labelled as:  

CUM6_BC_NO2_PM  a cluster of BC, NO2, and PM  
CUM6_PAH36  a large cluster of 36 PAHs  
CUM6_DlP  DlP only  
CUM6_NkF  NkF only  
CUM6_RET  retene only  
CUM6_SO2  SO2 only   
 

## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusCUM6, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUM6_mix, fig.width=8, fig.height=7}
rlts_gee_3 <- perform_group_gee(X = var_name_list$cluster_exp$clusCUM6, adjust_cfd = T)

for(j in 1:length(rlts_gee_3[[2]])){
  rlts_gee_3[[2]][[j]] <- rlts_gee_3[[2]][[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = rlts_gee_3[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee_3[[1]])){
  rlts_gee_3[[1]][[j]] <- rlts_gee_3[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee_3[[1]])
```



### GEE (with confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_NO2\_PM +  \beta_2 PAH36 + \beta_3 DlP  + \beta_4 NkF + \beta_5 RET + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUM6_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUM6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

rlts_gee_3_mutually_adj <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(rlts_gee_3_mutually_adj) <- var_name_list$cluster_exp$clusCUM6

p1 <- make_forest_plot_lg(coef1, "CUM6_BC_NO2_PM")
p2 <- make_forest_plot_lg(coef2, "CUM6_PAH36")
p3 <- make_forest_plot_lg(coef3, "CUM6_DlP")
p4 <- make_forest_plot_lg(coef4, "CUM6_NkF")
p5 <- make_forest_plot_lg(coef5, "CUM6_RET")
p6 <- make_forest_plot_lg(coef6, "CUM6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUM6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```


## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusCUM6_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusCUM6, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee[[1]])
```



### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_NO2\_PM +  \beta_2 PAH36 + \beta_3 DlP  + \beta_4 NkF + \beta_5 RET + \beta_6 SO2  \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusCUM6_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusCUM6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "CUM6_BC_NO2_PM")
p2 <- make_forest_plot_lg(coef2, "CUM6_PAH36")
p3 <- make_forest_plot_lg(coef3, "CUM6_DlP")
p4 <- make_forest_plot_lg(coef4, "CUM6_NkF")
p5 <- make_forest_plot_lg(coef5, "CUM6_RET")
p6 <- make_forest_plot_lg(coef6, "CUM6_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-3.5, 3.5))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusCUM6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```




# 3.0. All plots together
## Not mutually adjusted
```{r fig.width=15, fig.height=8}
rlts_gee_1[[1]]$CUR6_BC_PAH6$cluster = "BC_PAH6"
rlts_gee_1[[1]]$CUR6_PAH31$cluster = "PAH31"
rlts_gee_1[[1]]$CUR6_NkF$cluster = "NkF"
rlts_gee_1[[1]]$CUR6_PM_RET$cluster = "PM_RET"
rlts_gee_1[[1]]$CUR6_NO2$cluster = "NO2"
rlts_gee_1[[1]]$CUR6_SO2$cluster = "SO2"

rlts_gee_1_Df <- rbind(rlts_gee_1[[1]]$CUR6_BC_PAH6, rlts_gee_1[[1]]$CUR6_PAH31, 
                       rlts_gee_1[[1]]$CUR6_NkF, rlts_gee_1[[1]]$CUR6_PM_RET,
                       rlts_gee_1[[1]]$CUR6_NO2, rlts_gee_1[[1]]$CUR6_SO2)
rlts_gee_1_Df$exposure_period = "Current Exposure"

rlts_gee_2[[1]]$CHLD5_X7$cluster = "X7"
rlts_gee_2[[1]]$CHLD5_X33$cluster = "X33"
rlts_gee_2[[1]]$CHLD5_NkF$cluster = "NkF"
rlts_gee_2[[1]]$CHLD5_NO2$cluster = "NO2"
rlts_gee_2[[1]]$CHLD5_SO2$cluster = "SO2"
empty_df <- as.data.frame(matrix(NA, nrow = 8, ncol = 6)) 
empty_df$cluster = " "
row.names(empty_df) <- row.names(rlts_gee_2[[1]]$CHLD5_X7)
colnames(empty_df) <- colnames(rlts_gee_2[[1]]$CHLD5_X7)

rlts_gee_2_Df <- rbind(rlts_gee_2[[1]]$CHLD5_X7, rlts_gee_2[[1]]$CHLD5_X33,
                       rlts_gee_2[[1]]$CHLD5_NkF, rlts_gee_2[[1]]$CHLD5_NO2,
                       rlts_gee_2[[1]]$CHLD5_SO2, empty_df)
rlts_gee_2_Df$exposure_period = "Childhood Exposure"


rlts_gee_3[[1]]$CUM6_BC_NO2_PM$cluster = "BC_NO2_PM"
rlts_gee_3[[1]]$CUM6_PAH36$cluster = "PAH36"
rlts_gee_3[[1]]$CUM6_DlP$cluster = "DlP"
rlts_gee_3[[1]]$CUM6_NkF$cluster = "NkF"
rlts_gee_3[[1]]$CUM6_RET$cluster = "RET"
rlts_gee_3[[1]]$CUM6_SO2$cluster = "SO2"

rlts_gee_3_Df <- rbind(rlts_gee_3[[1]]$CUM6_BC_NO2_PM, rlts_gee_3[[1]]$CUM6_PAH36, 
                       rlts_gee_3[[1]]$CUM6_DlP, rlts_gee_3[[1]]$CUM6_NkF,
                       rlts_gee_3[[1]]$CUM6_RET, rlts_gee_3[[1]]$CUM6_SO2)
rlts_gee_3_Df$exposure_period = "Cumulative Exposure"

rlts_gee_Df <- rbind(rlts_gee_1_Df, rlts_gee_2_Df, rlts_gee_3_Df)


rlts_gee_Df$exposure_period = factor(rlts_gee_Df$exposure_period, 
                                     levels = c("Current Exposure","Childhood Exposure", "Cumulative Exposure"))
rlts_gee_Df$cluster = factor(rlts_gee_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
# rlts_gee_Df$cluster = factor(rlts_gee_Df$cluster, 
#                                      levels = c("BC_NO2_PM", "PAH36", "DlP", "BC_PAH6", "PAH31", "X7", "X33", 
#                                                 "PM_RET","RET",  "NkF", "NO2", "SO2", " "))
rlts_gee_Df$EAAs <- rep(var_name_list$EEAs, 6+6+6)
```


```{r gee_clusters_3periods, fig.width=15, fig.height=8}
make_forest_plot_lg(rlts_gee_Df, "Expected EAA Change with 1 Unit Increased in each Cluster") + 
  facet_wrap( ~ exposure_period + cluster, nrow=3) +
  ylim(-1.7,2.2)
```

```{r gee_clusters_3periods_sep, fig.width=15, fig.height=8}
rlts_gee_1_Df$cluster = factor(rlts_gee_1_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
rlts_gee_1_Df$EAAs <- rep(var_name_list$EEAs, 6)

rlts_gee_2_Df$cluster = factor(rlts_gee_2_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
rlts_gee_2_Df$EAAs <- rep(var_name_list$EEAs, 6)

rlts_gee_3_Df$cluster = factor(rlts_gee_3_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
rlts_gee_3_Df$EAAs <- rep(var_name_list$EEAs, 6)

p1 <- make_forest_plot_lg(rlts_gee_1_Df, "") + 
  facet_grid( ~ cluster) +
  ylim(-1.7,2.2) +
  labs(y=" ")
ggsave(file=sprintf("%s/gee_clusters_current.png",plotFolder), p1, width = 15, height = 3, dpi = 1200) 

p2 <- make_forest_plot_lg(rlts_gee_2_Df, "") + 
  facet_grid( ~ cluster) +
  ylim(-1.7,2.2)  +
  labs(y=" ")
ggsave(file=sprintf("%s/gee_clusters_childhood.png",plotFolder), p2, width = 15, height = 3, dpi = 1200) 

p3 <- make_forest_plot_lg(rlts_gee_3_Df, "") + 
  facet_grid( ~ cluster) +
  ylim(-1.7,2.2)
ggsave(file=sprintf("%s/gee_clusters_cumulatife.png",plotFolder), p3, width = 15, height = 3, dpi = 1200) 
```



## Mutually adjusted
```{r fig.width=15, fig.height=8}
rlts_gee_1_mutually_adj$CUR6_BC_PAH6$cluster = "BC_PAH6"
rlts_gee_1_mutually_adj$CUR6_PAH31$cluster = "PAH31"
rlts_gee_1_mutually_adj$CUR6_NkF$cluster = "NkF"
rlts_gee_1_mutually_adj$CUR6_PM_RET$cluster = "PM_RET"
rlts_gee_1_mutually_adj$CUR6_NO2$cluster = "NO2"
rlts_gee_1_mutually_adj$CUR6_SO2$cluster = "SO2"

rlts_gee_1_Df <- rbind(rlts_gee_1_mutually_adj$CUR6_BC_PAH6, rlts_gee_1_mutually_adj$CUR6_PAH31, 
                       rlts_gee_1_mutually_adj$CUR6_NkF, rlts_gee_1_mutually_adj$CUR6_PM_RET,
                       rlts_gee_1_mutually_adj$CUR6_NO2, rlts_gee_1_mutually_adj$CUR6_SO2)
rlts_gee_1_Df$exposure_period = "Current Exposure"

rlts_gee_2_mutually_adj$CHLD5_X7$cluster = "X7"
rlts_gee_2_mutually_adj$CHLD5_X33$cluster = "X33"
rlts_gee_2_mutually_adj$CHLD5_NkF$cluster = "NkF"
rlts_gee_2_mutually_adj$CHLD5_NO2$cluster = "NO2"
rlts_gee_2_mutually_adj$CHLD5_SO2$cluster = "SO2"
empty_df <- as.data.frame(matrix(NA, nrow = 8, ncol = 7)) 
empty_df$cluster = " "
row.names(empty_df) <- row.names(rlts_gee_2_mutually_adj$CHLD5_X7)
colnames(empty_df) <- colnames(rlts_gee_2_mutually_adj$CHLD5_X7)

rlts_gee_2_Df <- rbind(rlts_gee_2_mutually_adj$CHLD5_X7, rlts_gee_2_mutually_adj$CHLD5_X33,
                       rlts_gee_2_mutually_adj$CHLD5_NkF, rlts_gee_2_mutually_adj$CHLD5_NO2,
                       rlts_gee_2_mutually_adj$CHLD5_SO2, empty_df)
rlts_gee_2_Df$exposure_period = "Childhood Exposure"


rlts_gee_3_mutually_adj$CUM6_BC_NO2_PM$cluster = "BC_NO2_PM"
rlts_gee_3_mutually_adj$CUM6_PAH36$cluster = "PAH36"
rlts_gee_3_mutually_adj$CUM6_DlP$cluster = "DlP"
rlts_gee_3_mutually_adj$CUM6_NkF$cluster = "NkF"
rlts_gee_3_mutually_adj$CUM6_RET$cluster = "RET"
rlts_gee_3_mutually_adj$CUM6_SO2$cluster = "SO2"

rlts_gee_3_Df <- rbind(rlts_gee_3_mutually_adj$CUM6_BC_NO2_PM, rlts_gee_3_mutually_adj$CUM6_PAH36, 
                       rlts_gee_3_mutually_adj$CUM6_DlP, rlts_gee_3_mutually_adj$CUM6_NkF,
                       rlts_gee_3_mutually_adj$CUM6_RET, rlts_gee_3_mutually_adj$CUM6_SO2)
rlts_gee_3_Df$exposure_period = "Cumulative Exposure"

rlts_gee_Df <- rbind(rlts_gee_1_Df, 
                     rlts_gee_2_Df %>% select(- EAAs), 
                     rlts_gee_3_Df %>% select(- EAAs))


rlts_gee_Df$exposure_period = factor(rlts_gee_Df$exposure_period, 
                                     levels = c("Current Exposure","Childhood Exposure", "Cumulative Exposure"))
rlts_gee_Df$cluster = factor(rlts_gee_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
rlts_gee_Df$EAAs <- rep(var_name_list$EEAs, 6+6+6)
```


```{r gee_clusters_3periods_mutually_adj, fig.width=15, fig.height=8}
make_forest_plot_lg(rlts_gee_Df, "Expected EAA Change with 1 Unit Increased in each Cluster") + 
  facet_wrap( ~ exposure_period + cluster, nrow=3)

```

```{r gee_clusters_3periods_sep_mutually_adj, fig.width=15, fig.height=8}
rlts_gee_1_Df$cluster = factor(rlts_gee_1_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
rlts_gee_1_Df$EAAs <- rep(var_name_list$EEAs, 6)

rlts_gee_2_Df$cluster = factor(rlts_gee_2_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
rlts_gee_2_Df$EAAs <- rep(var_name_list$EEAs, 6)

rlts_gee_3_Df$cluster = factor(rlts_gee_3_Df$cluster, 
                                     levels = c("BC_NO2_PM", "PAH36",  "BC_PAH6", "PAH31", "X7", "X33", 
                                                "NkF",  "SO2","NO2", "RET", " ","DlP", "PM_RET"))
rlts_gee_3_Df$EAAs <- rep(var_name_list$EEAs, 6)

p1 <- make_forest_plot_lg(rlts_gee_1_Df, "") + 
  facet_grid( ~ cluster) +
  ylim(-1.7,2.2) +
  labs(y=" ")
ggsave(file=sprintf("%s/gee_clusters_current_mutual_adj.png",plotFolder), p1, width = 15, height = 3, dpi = 1200) 

p2 <- make_forest_plot_lg(rlts_gee_2_Df, "") + 
  facet_grid( ~ cluster) +
  ylim(-1.7,2.2)  +
  labs(y=" ")
ggsave(file=sprintf("%s/gee_clusters_childhood_mutual_adj.png",plotFolder), p2, width = 15, height = 3, dpi = 1200) 

p3 <- make_forest_plot_lg(rlts_gee_3_Df, "") + 
  facet_grid( ~ cluster) +
  ylim(-1.7,2.2)
ggsave(file=sprintf("%s/gee_clusters_cumulatife_mutual_adj.png",plotFolder), p3, width = 15, height = 3, dpi = 1200) 
```




# 3.4. Clusters based on pollutant measurements (clusMEAS6)
The file LEX_clusMEAS6.csv has information on measured pollutant exposures during each visit. Estimates are available for 6 different prototypes (cluster variables) for a total of 54 subjects and 54 visits. The prototypes are labelled as:   

MEAS6_BC_ PM_RET  a cluster of BC, PM, and retene  
MEAS6_X31  a large cluster of 31 air pollutants  
MEAS6_X5  a smaller cluster of 5 air pollutants  
MEAS6_DlP  DlP only  
MEAS6_NkF  NkF only  
MEAS6_ NO2_SO2  NO2, and SO2    


## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusMEAS6, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusMEAS6_mix, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusMEAS6, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee[[1]])
```



### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusMEAS6_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusMEAS6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "MEAS6_BC_PM_RET")
p2 <- make_forest_plot_lg(coef2, "MEAS6_X31")
p3 <- make_forest_plot_lg(coef3, "MEAS6_X5")
p4 <- make_forest_plot_lg(coef4, "MEAS6_DlP")
p5 <- make_forest_plot_lg(coef5, "MEAS6_NkF")
p6 <- make_forest_plot_lg(coef6, "MEAS6_NO2_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusMEAS6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```


## Sensitivity analyses
### GEE (no confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusMEAS6_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusMEAS6, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee[[1]])
```




### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * BC\_PAH6 +  \beta_2 PAH31 + \beta_3 NkF + \beta_4 PM\_RET + \beta_5 NO2 + \beta_6 SO2  \\
  & + \beta_7 * county + \beta_8 * BMI + \beta_9 * ses + \beta_10 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusMEAS6_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))
coef6 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusMEAS6,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
    coef6[i,] <- rlts[7,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)
colnames(coef6) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs
rownames(coef6) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef6$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "MEAS6_BC_PM_RET")
p2 <- make_forest_plot_lg(coef2, "MEAS6_X31")
p3 <- make_forest_plot_lg(coef3, "MEAS6_X5")
p4 <- make_forest_plot_lg(coef4, "MEAS6_DlP")
p5 <- make_forest_plot_lg(coef5, "MEAS6_NkF")
p6 <- make_forest_plot_lg(coef6, "MEAS6_NO2_SO2")

plot_list <- list(p1, p2, p3, p4, p5, p6)

for(j in 1:6){
  plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured pollutant exposures"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5, coef6)
names(coefs_list) <- var_name_list$cluster_exp$clusMEAS6
cat("The estimated effects:")
for(j in 1:6){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```



# 3.5. Clusters based on urinary biomarkers (clusURI5)
The file LEX_clusURI5.csv has information on measured urinary biomarkers obtained during each visit. Estimates are available for 5 different prototypes (cluster variables) for a total of 163 subjects and 186 visits. The prototypes are labelled as:  


URI5_NAP_1M_2M  a cluster of Naphthalene, 1Methylnaphthalene, and 2Methylnaphthalene  
URI5_ACE  Acenaphthene only  
URI5_FLU_PHE  Fluoranthene and Phenanthrene_anth  
URI5_PYR  Pyrene only  
URI5_CHR  Baa_Chrysene only  
 



## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$cluster_exp$clusURI5, collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```

## Primary analysis
### GEE (with confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X   \\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusURI5_mix, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusURI5, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured urinary biomarkers"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee[[1]])
```



### GEE (Mix, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * NAP\_1M\_2M +  \beta_2 ACE + \beta_3 FLU\_PHE + \beta_4 PYR + \beta_5 CHR  \\
  & + \beta_6 * county + \beta_7 * BMI + \beta_8 * ses + \beta_9 * edu + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusURI5_mix_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusURI5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "URI5_NAP_1M_2M")
p2 <- make_forest_plot_lg(coef2, "URI5_ACE")
p3 <- make_forest_plot_lg(coef3, "URI5_FLU_PHE")
p4 <- make_forest_plot_lg(coef4, "URI5_PYR")
p5 <- make_forest_plot_lg(coef5, "URI5_CHR")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured urinary biomarkers"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusURI5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```


## Sensitivity analyses
### GEE (No confounders)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * X + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations and X is one of the cluster estimates. 


Results:
```{r gee_clusURI5_simple, fig.width=8, fig.height=7}
rlts_gee <- perform_group_gee(X = var_name_list$cluster_exp$clusURI5, adjust_cfd = T)

for(j in 1:length(rlts_gee[[2]])){
  rlts_gee[[2]][[j]] <- rlts_gee[[2]][[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = rlts_gee[[2]],  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on cumulative pollutant exposures"), gp=gpar(fontsize=15, fontface = 'bold')))
```
```{r}
cat("The estimated effects:")
for(j in 1:length(rlts_gee[[1]])){
  rlts_gee[[1]][[j]] <- rlts_gee[[1]][[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(rlts_gee[[1]])
```



### GEE (No confounders, mutual adjust)
In this section, we perform the generalized estimating equations (GEE) to evaluate the associations between each cluster within current pollutant exposures and each Epigenetic Age Acceleration with the formula: 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 * NAP\_1M\_2M +  \beta_2 ACE + \beta_3 FLU\_PHE + \beta_4 PYR + \beta_5 CHR  \\
  & + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.


Results:
```{r gee_clusURI5_simple_mutual_adjust, fig.width=8, fig.height=7}
coef1 <- data.frame(matrix(NA, ncol = 6))
coef2 <- data.frame(matrix(NA, ncol = 6))
coef3 <- data.frame(matrix(NA, ncol = 6))
coef4 <- data.frame(matrix(NA, ncol = 6))
coef5 <- data.frame(matrix(NA, ncol = 6))

for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee_exposure_prototype(X = var_name_list$cluster_exp$clusURI5,
                                          Y = var_name_list$EEAs[i],
                                          adjust_cfd = T)
    coef1[i,] <- rlts[2,]
    coef2[i,] <- rlts[3,]
    coef3[i,] <- rlts[4,]
    coef4[i,] <- rlts[5,]
    coef5[i,] <- rlts[6,]
}

colnames(coef1) = colnames(rlts)
colnames(coef2) = colnames(rlts)
colnames(coef3) = colnames(rlts)
colnames(coef4) = colnames(rlts)
colnames(coef5) = colnames(rlts)

rownames(coef1) = var_name_list$EEAs
rownames(coef2) = var_name_list$EEAs
rownames(coef3) = var_name_list$EEAs
rownames(coef4) = var_name_list$EEAs
rownames(coef5) = var_name_list$EEAs


coef1$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef2$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef3$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef4$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))
coef5$EAAs = factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

p1 <- make_forest_plot_lg(coef1, "URI5_NAP_1M_2M")
p2 <- make_forest_plot_lg(coef2, "URI5_ACE")
p3 <- make_forest_plot_lg(coef3, "URI5_FLU_PHE")
p4 <- make_forest_plot_lg(coef4, "URI5_PYR")
p5 <- make_forest_plot_lg(coef5, "URI5_CHR")

plot_list <- list(p1, p2, p3, p4, p5)

for(j in 1:5){
 plot_list[[j]] <- plot_list[[j]] + ylim(c(-6, 6))
}

grid.arrange(grobs = plot_list,  ncol = 2,
             top = textGrob(paste("Estimated Effect of each Exposure Prototype \n based on measured urinary biomarkers"), 
                            gp=gpar(fontsize=15, fontface = 'bold')))

```

```{r}
coefs_list <- list(coef1, coef2, coef3, coef4, coef5)
names(coefs_list) <- var_name_list$cluster_exp$clusURI5
cat("The estimated effects:")
for(j in 1:5){
  coefs_list[[j]] <- coefs_list[[j]][, -7] %>%
    mutate(coefficient = round(coefficient, 4),                        
             std = round(as.numeric(std), 4),                        
             ci_lower = round(ci_lower, 4),                        
             ci_upper = round(ci_upper, 4),                        
             p_val = round(as.numeric(p_val), 6))
}
print(coefs_list)
```



# 4.1. Current exposure to 5MC
## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC[1], collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```
## Primary analysis
### GEE (with confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cur\_5mc\\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cur_5MC, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cur_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Current exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)
plot_df_1 <- plot_df
```





## Sensitivity analysis
### GEE (no confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cur\_5mc + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cur_5MC_simple, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cur_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Current exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)
```





# 4.2. Childhood exposure to 5MC
## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC[3], collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```
## Primary analysis

### GEE (with confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *childhood\_5mc\\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_bir_5MC, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "bir_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Childhood exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)
plot_df_2 <- plot_df
```





## Sensitivity analysis
### GEE (no confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *childhood\_5mc + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cir_5MC_simple, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cur_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Childhood exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)

```



# 4.3. Cumulative exposure to 5MC
## Summary the exposure estimates:
```{r}
table1(as.formula(paste(" ~ ", paste(var_name_list$PCH_5MC[2], collapse = " + "), "| curFuel")), 
       data = df,
       render.continuous=c(.="Mean (SD)", 
                           .="Median (IQR)"))
```
## Primary analysis

### GEE (with confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cum\_5mc\\
  & + \beta_2 * county + \beta_3 * BMI + \beta_4 * ses + \beta_5 * edu  + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cum_5MC, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cum_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Cumulatice exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)

plot_df_3 <- plot_df
```





## Sensitivity analysis
### GEE (no confounders)
In the following section, we performed generalized estimating equations (GEE) with equation 
$$
\begin{aligned}
  Y = & \beta_0 + \beta_1 *cum\_5mc + \epsilon
\end{aligned}
$$
where $Y$ is one of the epigenetic age accelerations.

Results:
```{r gee_coef_cum_5MC_simple, fig.width=4, fig.height=3, warning=FALSE, message=FALSE}
  lm_coef_ambient <- c()
  plot_df <- data.frame(matrix(NA, ncol = 6))
  for(i in 1:length(var_name_list$EEAs)){
    rlts = preform_gee(Y = var_name_list$EEAs[i], 
                              X = "cum_5mc", 
                              adjust_cfd = T, 
                              dat=df)
    lm_coef_ambient[i] <- rlts[1]
    plot_df[i,] <- rlts
  }
  colnames(plot_df) <- c("coefficient", "std", "ci_lower", "ci_upper", "p_val", "sig_level")
  rownames(plot_df) <- var_name_list$EEAs
  plot_df$EAAs <- EAAs_names
  plot_df$EAAs <- factor(EAAs_names, levels=rev(c(EAAs_names[EAAs_names != "DNAmTL"], "DNAmTL")))

  plot_df$coefficient = as.numeric(plot_df$coefficient)
  plot_df$ci_lower = as.numeric(plot_df$ci_lower)
  plot_df$ci_upper = as.numeric(plot_df$ci_upper)

  lm_coef_ambient_plot <- make_forest_plot(plot_df, "Current exposure to 5MC")

lm_coef_ambient_plot

print(" Result data: ")
print(plot_df)

```



# 4.0. All plots together
```{r gee_5MC_3Periods, fig.width=8, fig.height=3}
plot_df_1$exposure_time = "Current Exposure"
plot_df_2$exposure_time = "Childhood Exposure"
plot_df_3$exposure_time = "Cumulative Exposure"
plot_df_5mc <- rbind(plot_df_1, plot_df_2, plot_df_3)
plot_df_5mc$exposure_time = factor(plot_df_5mc$exposure_time, levels = c("Current Exposure","Childhood Exposure", "Cumulative Exposure"))
make_forest_plot_lg(plot_df_5mc, "Expected EAA Change with 1 Unit Increased in 5MC Exposure") + 
  facet_wrap(~ exposure_time, scales = 'free_x') 
```

```{r}
p1 <- make_forest_plot_lg(plot_df_1, "") + 
  labs(y = " ")
ggsave(file=sprintf("%s/5MC_current.png",plotFolder), p1, width = 4.5, height = 3, dpi = 1200) 

p2 <- make_forest_plot_lg(plot_df_2, "") 
ggsave(file=sprintf("%s/5MC_childhood.png",plotFolder), p2, width = 4.5, height = 3, dpi = 1200) 

p3 <- make_forest_plot_lg(plot_df_3, "") + 
  labs(y = " ")
ggsave(file=sprintf("%s/5MC_cumulative.png",plotFolder), p3, width = 4.5, height = 3, dpi = 1200) 

```
